### 2006-12-09 inst/doc/filtering-flowCore.Rnw
BIG set of changes! This changes the implementation of the "filter" method by moving most of the work out of the functions and into the %in% method, which is exposed to the user. This is designed to be easier for programmers and future filter implementors.
filter() itself is now a single function that dispatches the appropriate %in%. Also, we're now carrying around some extra filterResult data to help us make filterResult a type of filter class that can be used in conjunction with other unresolved filters or with sets of filter results using the usual sorts of logical operators. 

We also have

% NOTE -- ONLY EDIT THE .Rnw FILE!!!  The .tex file is
% likely to be overwritten.
%
%\VignetteIndexEntry{Filtering Flow Cytometry Data}
%\VignetteDepends{flowCore}
%\VignetteKeywords{}
%\VignettePackage{flowCore}
\documentclass[11pt]{article}

\usepackage{times}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{times}
\usepackage{comment}
\usepackage{graphicx}
\usepackage{subfigure}

\textwidth=6.2in
\textheight=8.5in
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

\newcommand{\scscst}{\scriptscriptstyle}
\newcommand{\scst}{\scriptstyle}
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textsf{#1}}}
\newcommand{\Rclass}[1]{{\textit{#1}}}

\bibliographystyle{plainnat}

\title{Filtering in the flowCore package}
\author{N. Le Meur F. Hahne B. Ellis P. Haaland}

\begin{document}
\maketitle

\begin{abstract}
\noindent One of the more common operations in flow cytometry data analysis is a process of filtering, also known as gating. This is often either the target of analysis in the case of surface staining or a necessary pre-processing step for further analysis on the cell population of interest. For example, a researcher may be interested in CD4+ T cells, which typically requires a lymphocyte gate and a CD4 gate to obtain a relatively clean population and other cell types might be even more complex. In the case where the number of cells within a gate is of interest, it is not uncommon to require a fairly complex composition of gates for analysis. This document describes the rich filter construction and composition facilities in the flowCore package as well as what is needed for developing your own filter objects. \\ \\
\textbf{keywords} Flow cytometry, high throughtput, software, standard
\end{abstract}

\end{document}