### 2007-05-30 inst/doc/HowTo-flowCore.Rnw
Add splitScaleTransformation (for transformML) Modify read.FCS to remove random sampling but leave which.lines Add parentId argument in class Filter Fix rectangle gate boundary allow Min>=Max like in GateML standard, return empty gate
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@25068 bc3139a8-67e5-0310-9ffc-ced21a209358

% NOTE -- ONLY EDIT THE .Rnw FILE!!!  The .tex file is
% likely to be overwritten.
%
%\VignetteIndexEntry{Basic Functions for Flow Cytometry Data}
%\VignetteDepends{flowCore}
%\VignetteKeywords{}
%\VignettePackage{flowCore}
\documentclass[11pt]{article}

\usepackage{times}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{times}
\usepackage{comment}
\usepackage{graphicx}
\usepackage{subfigure}

\textwidth=6.2in
\textheight=8.5in
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

\newcommand{\scscst}{\scriptscriptstyle}
\newcommand{\scst}{\scriptstyle}
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Rcode}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textsf{#1}}}
\newcommand{\Rclass}[1]{{\textit{#1}}}

\title{flowCore: data structures package for flow cytometry data}
\author{N. Le Meur F. Hahne B. Ellis P. Haaland}

\begin{document}
\maketitle

\begin{abstract}
\noindent \textbf{Background} The recent application of modern automation technologies to staining and collecting flow cytometry (FCM) samples has led to many new challenges in data management and analysis. We limit our attention here to the associated problems in the analysis of the massive amounts of FCM data now being collected. From our viewpoint, see two related but substantially different problems arising. On the one hand, there is the problem of adapting existing software to apply standard methods to the increased volume of data. The second problem, which we intend to address here, is the absence of any research platform which bioinformaticians, computer scientists, and statisticians can use to develop novel methods that address both the volume and multidimensionality of the mounting tide of data. In our opinion, such a platform should be Open Source, be focused on visualization,  support rapid prototyping, have a large existing base of users, and have demonstrated suitability for development of new methods. We believe that the Open Source statistical software R in conjunction with the Bioconductor Project fills all of these requirements. Consequently we have developed a Bioconductor package that we call \Rpackage{flowCore}. The \Rpackage{flowCore} package is not intended to be a complete analysis package for FCM data, rather, we see it as providing a clear object model and a collection of standard tools that enable R as an informatics research platform for flow cytometry. One of the important issues that we have addressed in the \Rpackage{flowCore} package is that of using a standardized representation that will insure compatibility with existing technologies for data analysis and will support collaboration and interoperability of new methods as they are developed. In order to do this, we have followed the current standardized descriptions of FCM data analysis as being developed under NIH Grant xxxx [n]. We believe that researchers will find flowCore to be a solid foundation for future development of new methods to attack the many interesting open research questions in FCM data analysis.\\

\noindent \textbf{Methods} We propose a variety different data structures. We have implemented the classes and methods in the Bioconductor package flowCore. 
We illustrate their use with X case studies.\\

\noindent \textbf{Results} We hope that those proposed data structures will the base for the development of many tools for the analysis of high throughput flowcytometry.\\

\noindent \textbf{keywords} Flow cytometry, high throughtput, software, standard
\end{abstract}

<<loadPackage, echo=false,results=hide>>=
library(flowCore)
@

\section{Introduction}
Traditionally, flow cytometry has been a tube-based technique limited to small-scale laboratory and clinical studies. 
High throughput methods for flow cytometry have recently been developed for drug discovery and advanced research methods 
\citep{Gasparetto2004}. As an example, the flow cytometry high content screening (FC-HCS) can process up 
to a thousand samples daily at a single workstation, and the results have been equivalent or superior to traditional 
manual multiparameter staining and analysis techniques. \\

The amount of information generated by high throughput technologies such as FC-HCS need to be transformed into 
executive summaries (which are brief enough) for creative studies by a human researcher \citep{Brazma2001}. 
Standardization is critical when developing new high throughput technologies and their associated information services 
\citep{Brazma2001, Chicurel2002, Boguski2003}. Standardization efforts have been made in clinical cell analysis 
by flow cytometry \citep{Keeney2004}, however data interpretation has not been standardized for even low throughput FCM. 
It is one of the most difficult and time consuming aspects of the entire analytical process as well as a primary source 
of variation in clinical tests, and investigators have traditionally relied on intuition rather than standardized statistical inference 
\citep{Bagwell2004,Braylan2004,Parks1997,Suni2003}. In the development of standards in high throughput FCM, 
few progress has been done in term of Open Source software. In this article we propose R data structures to handle 
flow cytometry data through the main steps of preprocessing: compensation, transformation, filtering.\\

The aim is to merge both prada  and rflowcyt \citep{LeMeur2006} into one core package wich is compliant 
with the data exchange standards that are currently developed in the community \citep{Spidlen2006}.

Visualization as well as quality control will than be part of the utility packages that depend on the data structures defined in the flowCore package.

\section{Representing Flow Cytometry Data}

\Rpackage{flowCore}'s primary task is the representation and basic manipulation of flow cytometry (or similar) data. This is accomplished through a data model very similar to that adopted by other Bioconductor packages using the \Robject{expressionSet} and \Robject{AnnotatedDataFrame} structures familiar to most Bioconductor users.

\subsection{The \Rclass{flowFrame} Class}

The basic unit of manipulation in \Rpackage{flowCore} is the
\Rclass{flowFrame}, which corresponds roughly with a single
``FCS'' file exported from the flow cytometer's acquisition
software. At the moment we support FCS file versions 1.0
through 3.0, and we expect to support FCS4/ACS1 as soon as
the specification has been ratified.

\subsubsection{Data elements}

The primary elements of the \Robject{flowFrame} are the
\Robject{exprs} and \Robject{parameters} slots, which
contain the event-level information and column metadata
respectively. The event information, stored as a single
matrix, is accessed and manipulated via the
\Rfunction{exprs()} and \Rfunction{exprs<-} methods,
allowing \Robject{flowFrame}s to be stitched together if
necessary (for example, if the same tube has been collected
in two acquisition files for memory reasons).

The \Robject{parameters} slot is an
\Rclass{AnnotatedDataFrame} that contains information
derived from an FCS file's ``\$P<n>'' keywords, which
describe the detector and stain information. The entire list
is available via the \Rfunction{parameter()} method, but
more commonly this information is accessed through the
\Rfunction{names}, \Rfunction{featureNames} and
\Rfunction{colnames} methods. The \Rfunction{names} function
returns a concatenated version of \Rfunction{names} and
\Rfunction{featureNames} using a format similar to the one
employed by most flow cytometry analysis software. The
\Rfunction{colnames} method returns the detector names,
often named for the fluorochrome detected, while the
\Rfunction{featureNames} methods returns the description
field of the parameters, which will typically be an
identifier for the antibody.

The \Rfunction{keyword} method allows access to the raw FCS
keywords, which are a mix of standard entries such as
``SAMPLE ID,'' vendor specific keywords and user-defined
keywords that add more information about an experiment. In
the case of plate-based experiments, there are also one or
more keywords that identify the specific well on the plate.

Most vendor software also include some sort of unique
identifier for the file itself. The specialized methods
\Rfunction{identifier} attempts to locate an appropriate
globally unique identifier that can be used to uniquely
identify a frame. Failing that, this method will return the
original file name offering some assurance that this frame
is at least unique to a particular session.

\subsubsection{Reading a \Robject{flowFrame}}

FCS files are read into the R environment via the
\Rfunction{read.FCS} function using the standard connection
interface---allowing for the possibility of accessing FCS
files hosted on a remote resource as well as those that have
been compressed or even retrieved as a blob from a database
interface. FCS files (version 2.0 and 3.0) and LMD (List
Mode Data) extensions are currently supported.

There are also several immediate processing options
available in this function, the most important of which is
the \Rfunction{transformation} parameter, which can either
``linearize'' (the default) or ``standardize'' our data. To
see how this works, first we will examine an FCS file
without any transformation at all:

<<ReadFiles, echo=true, results=verbatim>>=
file.name <- system.file("extdata","0877408774.B08", package="flowCore")
x <- read.FCS(file.name, transformation=FALSE)
summary(x)
@

As we can see, in this case the values from each parameter
seem to run from $0$ to $1023$ ($2^{10}-1$).  However,
inspection of the ``exponentiation'' keyword (\$P<n>E)
reveals that some of the parameters (3 and 4) have been
stored in the format of the form $a\times{10^{x/R}}$ where
$a$ is given by the first element of the string.

<<SearchKeywords, echo=true,results=verbatim>>=
keyword(x,c("$P1E", "$P2E", "$P3E", "$P4E"))
@

The default ``linearize'' transformation option will convert these to, effectively, have a ``\$P<n>E'' of ``0,0'':

<<PrintSummary, echo=true, results=verbatim>>=
summary(read.FCS(file.name))
@

Finally, the ``scale'' option will both linearize values as
well as ensure that output values are contained in $[0,1]$,
which is the proposed method of data storage for the
ACS1.0/FCS4.0 specification:

<<PrintSummary2, echo=true,results=verbatim>>=
summary(read.FCS(file.name,transformation="scale")) 
@

Another parameter of interest is the \Rfunction{alter.names}
parameter, which will convert the parameter names into more
``R friendly'' equivalents, usually by replacing ``-'' with
``.'':

<<ReadFiles2, echo=true,results=verbatim>>=
read.FCS(file.name,alter.names=TRUE) 
@

When only a particular subset of parameters is
desired the \Rfunction{column.pattern} parameter allows for
the specification of a regular expression and only
parameters that match the regular expression will be
included in the frame. For example, to include on the the
Height parameters:

<<RedFiles3, echo=true, results=verbatim>>= 
x <- read.FCS(file.name, column.pattern="-H") 
x 
@

\textit{Note that \Rfunction{column.pattern} is applied after \Rfunction{alter.names} if it is used.}

Finally, only a sample of lines can be read in case you need a quick overview of a large serie of files. 
<<RedFiles3, echo=true, results=verbatim>>= 
lines <- sample(100:500, 50)
y <- read.FCS(file.name, which.lines = lines) 
y 
@

\subsubsection{Visualizing a \Rclass{flowFrame}}

Much of the more sophisticated visualization of \Rclass{flowFrame} and \Rclass{flowSet} objects, including an interface to the \Rpackage{lattice} graphics system is implemented by the \Rpackage{flowViz} package, also included as part of Bioconductor. However, \Rpackage{flowCore} does implement some basic plotting facilities using the standard \Rfunction{plot} function. The basic plot provides a simple bivariate density plot of the first two parameters:

<<Plot1, fig=true, echo=true>>=
plot(x)
@

To control the parameters being plotted we can supply a \Robject{y} value in the form of a character vector:

<<Plot2, fig=true, echo=true>>=
plot(x,c("FL1-H", "FL2-H"))
@

However, if we only supply a single parameter we instead get
a univariate histogram, which also accepts the usual
histogram arguments:

<<plot3, fig=true, echo=true>>=
plot(x, "FL1-H", breaks=256)
@

\subsection{The \Rclass{flowSet} Class}

Most experiments consist of several \Rclass{flowFrame} objects, which are organized using a \Rclass{flowSet} object. This class provides a mechanism for efficiently hosting the \Rclass{flowFrame} objects with minimal copying, reducing memory requirements, as well as ensuring that experimental metadata stays properly to the appropriate \Rclass{flowFrame} objects.

\subsubsection{Creating a \Rclass{flowSet}}

To facilitate the creation of \Rclass{flowSet} objects from a variety of sources, we provide a means to coerce \Rclass{list} and \Rclass{environment} objects to a \Rclass{flowSet} object using the usual coercion mechanisms. For example, if we have a directory containing FCS files we can read in a list of those files and create a \Rclass{flowSet} out of them:

<<Frames, echo=true, results=verbatim>>=
frames <- lapply(dir(system.file("extdata", "compdata", "data", package="flowCore"), full.names=TRUE), read.FCS)
as(frames, "flowSet")
@

Note that the original list is unnamed and that the resulting sample names are not particularly meaningful. If the list is named, the list constructed is much more meaningful. One such approach is to employ the \Rfunction{keyword} method for \Rclass{flowFrame} objects to extract the ``SAMPLE ID'' keyword from each frame:

<<Frames, echo=true,results=verbatim>>=
names(frames) <- sapply(frames, keyword, "SAMPLE ID")
fs <- as(frames, "flowSet")
fs
@

\subsubsection{Working with experimental metadata}

Like most Bioconductor organizational classes, the \Rclass{flowSet} has an associated \Rclass{AnnotatedDataFrame} that provides metadata not contained within the \Rclass{flowFrame} objects themselves. This data frame is accessed and modified via the usual \Rfunction{phenoData} and \Rfunction{phenoData<-} methods. You can also generally treat the phenotypic data as a normal data frame to add new descriptive columns. For example, we might want to track the original filename of the frames from above in the phenotypic data for easier access:

<<metaData, echo=true, results=verbatim>>=
phenoData(fs)$Filename <- fsApply(fs,keyword, "$FIL")
pData(phenoData(fs))
@

Note that we have used the \Rclass{flowSet}-specific iterator, \Rfunction{fsApply}, which acts much like \Rfunction{sapply} or \Rfunction{lapply}. Additionally, we should also note that the \Rfunction{phenoData} data frame \textbf{must} have row names that correspond to the original names used to create the \Rclass{flowSet}.

\subsubsection{Bringing it all together: \Rfunction{read.flowSet}}

Much of the functionality described above has been packaged into the \Rfunction{read.flowSet} convenience function. In it's simplest incarnation, this function takes a \Rfunction{path}, that defaults to the current working directory, and an optional \Rfunction{pattern} argument that allows only a subset of files contained within the working directory to be selected. For example, to read a \Rclass{flowSet} of the files read in by \Robject{frame} above:

<<ReadFlowSet, echo=true,results=verbatim>>=
read.flowSet(path = system.file("extdata", "compdata", "data", package="flowCore"))
@

\Rfunction{read.flowSet} will pass on additional arguments meant for the underlying \Rfunction{read.FCS} function, such as \Rfunction{alter.names} and \Rfunction{column.pattern}, but also supports several other interesting arguments for conducting initial processing:
\begin{description}
\item[files]{ An alternative to the \Rfunction{pattern} argument, you may also supply a vector of filenames to read.}
\item[name.keyword]{ Like the example in the previous section, you may specify a particular keyword to use in place of the filename when creating the \Rclass{flowSet}.}
\item[phenoData]{ If this is an \Rclass{AnnotatedDataFrame}, then this will be used in place of the data frame that is ordinarily created. Additionally, the row names of this object will be taken to be the filenames of the FCS files in the directory specified by \Robject{path}. This argument may also be a named list made up of a combination of \Robject{character} and \Robject{function} objects that specify a keyword to extract from the FCS file or a function to apply to each frame that will return a result. }
\end{description}

To recreate the \Rclass{flowSet} that we created by hand from the last section we can use \Rfunction{read.flowSet}s advanced functionality:

<<ReadFowSet2, echo=true,results=verbatim>>=
fs <- read.flowSet(path=system.file("extdata", "compdata", "data", package="flowCore"), name.keyword="SAMPLE ID", phenoData=list(name="SAMPLE ID", Filename="$FIL"))
fs
pData(phenoData(fs))
@

\subsubsection{Manipulating a \Rclass{flowSet}}

You can extract a \Rclass{flowFrame} from a \Rclass{flowSet} object in the usual way using the \Rfunction{[[} or \Rfunction{\$} extraction operators. On the other hand using he \Rfunction{[} extraction operator returns a new \Rclass{flowSet} by \textbf{copying} the environment. However, simply assigning the \Rclass{flowFrame} to a new variable will \textbf{not} copying the contained frames.

The primary iterator method for a \Rclass{flowSet} is the \Rfunction{fsApply} method, which works more-or-less like \Rfunction{sapply} or \Rfunction{lapply} with two extra options. The first argument, \Robject{simplify}, which defaults to \Robject{TRUE}, instructs \Rfunction{fsApply} to attempt to simplify it's results much in the same way as \Rfunction{sapply}. The primary difference is that if all of the return values of the iterator are \Rclass{flowFrame} objects, \Rfunction{fsApply} will create a new \Rclass{flowSet} object to hold them. The second argument, \Robject{use.exprs}, which defaults to \Robject{FALSE} instructs \Rfunction{fsApply} to pass the expression matrix of each frame rather than the \Rclass{flowFrame} object itself. This allows functions to operate directly on the intensity information without first having to extract it.

As an aid to this sort of operation we also introduce the \Rfunction{each\_row} and \Rfunction{each\_col} convenience functions that take the place of \Rfunction{apply} in the \Rfunction{fsApply} call. For example, if we wanted the median value of each parameter of each \Rclass{flowFrame} we might write:

<<fsApply1, echo=true, results=verbatim>>=
fsApply(fs, each_col, median)
@

which is equivalent to the less readable

<<fsApply2, echo=true,results=verbatim>>=
fsApply(fs,function(x) apply(x, 2, median), use.exprs=TRUE)
@

In this case, the \Robject{use.exprs} argument is not required in the first case because \Rfunction{each\_col} and \Rfunction{each\_row} are methods and have been defined to work on \Rclass{flowFrame} objects by first extracting the intensity data.

\section{Transformation}

\Rpackage{flowCore} features two methods of transforming parameters within a \Rclass{flowFrame}: inline and out-of-line. The inline method, discussed in the next section has been developed primarily to support filtering features and is strictly more limited than the out-of-line transformation method, which uses R's \Rfunction{transform} function to accomplish the filtering. Like the normal \Rfunction{transform} function, the \Rclass{flowFrame}is considered to be a data frame with columns named for parameters of the FCS file. For example, if we wished to plot our first \Rclass{flowFrame}'s first two fluorescence parameters on the log scale we might write:

<<Transfo1, fig=true, echo=true>>=
plot(transform(fs[[1]], `FL1-H`=log(`FL1-H`), `FL2-H`=log(`FL2-H`)), c("FL1-H","FL2-H"))
@

Like the usual \Rfunction{transform} function, we can also create new parameters based on the old parameters, without destroying the old

<<Transfo2, fig=true, prefix=FALSE>>=
plot(transform(fs[[1]], log.FL1.H=log(`FL1-H`), log.FL2.H=log(`FL2-H`)), c("log.FL1.H", "log.FL2.H"))
@

\subsection{Standard Transforms}

Though any function can be used as a transform in both the out-of-line and inline transformation techniques, \Rpackage{flowCore} provides a number of parameterized transform generators that correspond to the transforms commonly found in flow cytometry and defined in the Transformation Markup Language (Transformation-ML). Briefly, the predefined transforms are:

\begin{description}
\item[truncateTransform]{ $y = \left\{\begin{array}{ll}
a & x < a \\
x & x \geq a \\
\end{array}\right.$ }
\item[scaleTransform]{$ f(x) = \frac{x-a}{b-a}$}
\item[linearTransform]{ $ f(x) = a+bx$ }
\item[quadraticTransform]{ $ f(x) = ax^2 + bx + c$ }
\item[lnTransform]{$ f(x) = \log\left(x\right)\frac{r}{d} $}
\item[logTransform]{$ f(x) = \log_b\left(x\right)\frac{r}{d} $}
\item[biexponentialTransform]{ $ f^{-1}(x) = ae^{bx}-ce^{dx}+f$}
\item[logicleTransform]{ A special form of the biexponential transform with parameters selected by the data.}
\item[arcsinhTransform]{ $ f(x) = asinh\left(a+bx\right)+c$}
\end{description}

To use a standard transform, first we create a transform function via the constructors supplied by \Rpackage{flowCore}:

<<Transfo3, echo=true>>=
aTrans <- truncateTransform("truncate at 1", a=1)
aTrans
@
which we can then use on the parameter of interest in the usual way
<<Transfo4, echo=true,results=verbatim>>=
transform(fs,`FL1-H`=aTrans(`FL1-H`))
@

\section{Filtering}

The most common task in the analysis of flow cytometry data is usual some form of filtering operation, either to obtain summary statistics about the number of events that meet a certain criteria or to perform further analysis on a subset of the data.

\subsection{Standard Filters}

Most filtering operations are a composition of one or more common filtering operations. Like transformations, \Rpackage{flowCore} includes a number of built-in common flow cytometry filters.

The simplest of these filters are the geometric filters, which correspond to those typically found in interactive flow cytometry software:
\begin{description}
\item[rectangleGate]{Describes a cubic shape in one or more dimensions--a rectangle in one dimension is simply an interval gate.}
\item[polygonGate]{Describes an arbitrary two dimensional polygonal gate.}
\item[polytopeGate]{Describes a region that is the convex hull of the given points. This gate can exist in dimensions higher than $2$, unlike the \Robject{polygonGate}.}
\item[ellipsoidGate]{Describes an ellipsoidal region in two or more dimensions}
\end{description}

These gates are all described in more or less the same manner (see man pages for more details):

<<rectGate, echo=true, results=verbatim>>=
rectGate <- rectangleGate(filterId="Fluorescence Region", "FL1-H"=c(50, 100), "FL2-H"=c(50, 100))
@

We also include some data-driven filters not usually found in flow cytometry software:

\begin{description}
\item[norm2Filter]{ A robust method for finding a region that most resembles a bivariate Normal distribution. }
\item[kmeansFilter]{ Identifies populations based on a one dimensional k-means clustering operation. Allows the specification of \textbf{multiple} populations. }
\end{description}

\subsection{Count Statistics}

When we have constructed a filter, we can apply it in two basic ways. The first is to collect simple summary statistics on the number and proportion of events considered to be contained within the gate or filter. This is done using the \Rfunction{filter} method. The first step is to apply our filter to some data

<<echo=true,results=verbatim>>=
result = filter(fs[[1]],rectGate)
result
@

As we can see, we have returned a \Rclass{filterResult} object, which is in turn a filter allowing for reuse in, for example, subsetting operations. To obtain count and proportion statistics, we take the summary of this \Rclass{filterResult}, which returns a list of summary values:

<<Summary3, echo=true, results=verbatim>>=
summary(result)
summary(result)$n
summary(result)$true
summary(result)$p
@

A filter which contains multiple populations, such as the \Rclass{kmeansFilter}, can return a list of summary lists:

<<SummarFilter, echo=true, results=verbatim>>=
summary(filter(fs[[1]], kmeansFilter("myKMeans","FSC-H"=c("Low", "Medium", "High"))))
@

A filter may also be applied to an entire \Rclass{flowSet}, in which case it returns a list of \Rclass{filterResult} objects:

<<echo=true,results=verbatim>>=
filter(fs,rectGate)
@

\subsection{Subsetting}

To subset or split a \Rclass{flowFrame} or \Rclass{flowSet}, we use the \Rfunction{Subset} and \Rfunction{split} methods respectively. The first, \Rfunction{Subset}, behaves similarly to the standard R \Rfunction{subset} function, which unfortunately could not used. For example, recall from our initial plots of this data that the morphology parameters, Forward Scatter and Side Scatter contain a large more-or-less ellipse shaped population. If we wished to deal only with that population, we might use \Rfunction{Subset} along with a \Rclass{norm2Filter} object as follows:

<<Norm2Filter, echo=true, results=verbatim>>=
morphGate <- norm2Filter(filterId="MorphologyGate", "FSC-H", "SSC-H", scale=2)
smaller <- Subset(fs, morphGate)
fs[[1]]
smaller[[1]]
@

Notice how the \Robject{smaller} \Rclass{flowFrame} objects contain fewer events. Now imagine we wanted to use a \Rclass{kmeansFilter} as before to split our first fluorescence parameter into three populations. To do this we employ the \Rfunction{split} function:

<<Split, echo=true, results=verbatim>>=
split(smaller[[1]], kmeansFilter("myKMeans","FSC-H"=c("Low","Medium","High")))
@

or for an entire \Rclass{flowSet}

<<Split2, echo=true, results=verbatim>>=
split(smaller, kmeansFilter("myKMeans", "FSC-H"=c("Low", "Medium", "High")))
@

\subsection{Combining Filters}

Of course, most filtering operations consist of more than one gate. To combine gates and filters we use the standard R Boolean operators: \&, | and ! to construct an intersection, union and complement respectively:

<<CombineFilter, echo=true, results=verbatim>>=

rectGate & morphGate
rectGate | morphGate
!morphGate

@

we also introduce the notion of the subset operation, denoted by either \%subset\% or \%\&\%. This combination of two gates first performs a subsetting operation on the input \Rclass{flowFrame} using the right-hand filter and then applies the left-hand filter. For example,

<<Summary5, echo=true, results=verbatim>>=
summary(filter(smaller[[1]],rectGate %&% morphGate))
@

first calculates a subset based on the \Robject{morphGate} filter and then applies the \Robject{rectGate}.

\subsection{Transformation Filters}

Finally, it is sometimes desirable to construct a filter with respect to transformed parameters. To allow for this in our filtering constructs we introduce a special form of the \Rfunction{transform} method along with another filter combination operator \%on\%, which can be applied to both filters and \Rclass{flowFrame} or \Rclass{flowSet} objects. To specify our transform filter we must first construct a transform list using a simplified version of the \Rfunction{transform} function:

<<Transfo5, echo=true,results=verbatim>>=
tFilter <- transform("FL1-H"=log,"FL2-H"=log)
tFilter
@

Note that this version of the transform filter does not take parameters on the right-hand side--the functions can only take a single vector that is specified by the parameter on the left-hand side. In this case those parameters are ``FL1-H'' and ``FL2-H.'' The function also does not take a specific \Rclass{flowFrame} or \Rclass{flowSet} allowing us to use this with any appropriate data. We can then construct a filter with respect to the transform as follows:

<<TectGate3, echo=true, results=verbatim>>=
rect2 <- rectangleGate(filterId="Another Rect", "FL1-H"=c(1,2), "FL2-H"=c(2,3)) %on% tFilter
rect2
@

Additionally, we can use this construct directly on a \Rclass{flowFrame} or \Rclass{flowSet} by moving the transform to the left-hand side and placing the data on the right-hand side:

<<Plot6, fig=true,echo=true>>=
plot(tFilter %on% smaller[[1]],c("FL1-H","FL2-H"))
@

which has the same effect as the $\log$ transform used earlier.

\clearpage
\bibliographystyle{plainnat} 
\bibliography{cytoref}
\end{document}