### 2008-04-08 inst/doc/HowTo-flowCore.Rnw
making sure that the samplesare in the same order after splitting. Also added a couple of updates to the documentation
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@30903 bc3139a8-67e5-0310-9ffc-ced21a209358

% NOTE -- ONLY EDIT THE .Rnw FILE!!!  The .tex file is
% likely to be overwritten.
%
%\VignetteIndexEntry{Basic Functions for Flow Cytometry Data}
%\VignetteDepends{flowCore}
%\VignetteKeywords{}
%\VignettePackage{flowCore}
\documentclass[11pt]{article}

\usepackage{times}
\usepackage{hyperref}
\usepackage[authoryear,round]{natbib}
\usepackage{times}
\usepackage{comment}
\usepackage{graphicx}
\usepackage{subfigure}

\textwidth=6.2in
\textheight=8.5in
\oddsidemargin=.1in
\evensidemargin=.1in
\headheight=-.3in

\newcommand{\scscst}{\scriptscriptstyle}
\newcommand{\scst}{\scriptstyle}
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Rcode}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textsf{#1}}}
\newcommand{\Rclass}[1]{{\textit{#1}}}

\title{flowCore: data structures package for flow cytometry data}
\author{N. Le Meur F. Hahne B. Ellis P. Haaland}

\begin{document}
\maketitle

\begin{abstract}
\noindent \textbf{Background} The recent application of modern automation technologies to staining and collecting flow cytometry (FCM) samples has led to many new challenges in data management and analysis. We limit our attention here to the associated problems in the analysis of the massive amounts of FCM data now being collected. From our viewpoint, see two related but substantially different problems arising. On the one hand, there is the problem of adapting existing software to apply standard methods to the increased volume of data. The second problem, which we intend to address here, is the absence of any research platform which bioinformaticians, computer scientists, and statisticians can use to develop novel methods that address both the volume and multidimensionality of the mounting tide of data. In our opinion, such a platform should be Open Source, be focused on visualization,  support rapid prototyping, have a large existing base of users, and have demonstrated suitability for development of new methods. We believe that the Open Source statistical software R in conjunction with the Bioconductor Project fills all of these requirements. Consequently we have developed a Bioconductor package that we call \Rpackage{flowCore}. The \Rpackage{flowCore} package is not intended to be a complete analysis package for FCM data, rather, we see it as providing a clear object model and a collection of standard tools that enable R as an informatics research platform for flow cytometry. One of the important issues that we have addressed in the \Rpackage{flowCore} package is that of using a standardized representation that will insure compatibility with existing technologies for data analysis and will support collaboration and interoperability of new methods as they are developed. In order to do this, we have followed the current standardized descriptions of FCM data analysis as being developed under NIH Grant xxxx [n]. We believe that researchers will find flowCore to be a solid foundation for future development of new methods to attack the many interesting open research questions in FCM data analysis.\\

\noindent \textbf{Methods} We propose a variety different data structures. We have implemented the classes and methods in the Bioconductor package flowCore. 
We illustrate their use with X case studies.\\

\noindent \textbf{Results} We hope that those proposed data structures will the base for the development of many tools for the analysis of high throughput flowcytometry.\\

\noindent \textbf{keywords} Flow cytometry, high throughtput, software, standard
\end{abstract}

<<loadPackage, echo=false,results=hide>>=
library(flowCore)
@

\section{Introduction}
Traditionally, flow cytometry has been a tube-based technique limited
to small-scale laboratory and clinical studies.  High throughput
methods for flow cytometry have recently been developed for drug
discovery and advanced research methods \citep{Gasparetto2004}. As an
example, the flow cytometry high content screening (FC-HCS) can
process up to a thousand samples daily at a single workstation, and
the results have been equivalent or superior to traditional manual
multiparameter staining and analysis techniques.

The amount of information generated by high throughput technologies
such as FC-HCS need to be transformed into executive summaries (which
are brief enough) for creative studies by a human researcher
\citep{Brazma2001}.  Standardization is critical when developing new
high throughput technologies and their associated information services
\citep{Brazma2001, Chicurel2002, Boguski2003}. Standardization efforts
have been made in clinical cell analysis by flow cytometry
\citep{Keeney2004}, however data interpretation has not been
standardized for even low throughput FCM.  It is one of the most
difficult and time consuming aspects of the entire analytical process
as well as a primary source of variation in clinical tests, and
investigators have traditionally relied on intuition rather than
standardized statistical inference
\citep{Bagwell2004,Braylan2004,Parks1997,Suni2003}. In the development
of standards in high throughput FCM, few progress has been done in
term of Open Source software. In this article we propose R data
structures to handle flow cytometry data through the main steps of
preprocessing: compensation, transformation, filtering.

The aim is to merge both prada and rflowcyt \citep{LeMeur2006} into
one core package wich is compliant with the data exchange standards
that are currently developed in the community \citep{Spidlen2006}.

Visualization as well as quality control will than be part of the
utility packages that depend on the data structures defined in the
flowCore package.

\section{Representing Flow Cytometry Data}

\Rpackage{flowCore}'s primary task is the representation and basic
manipulation of flow cytometry (or similar) data. This is accomplished
through a data model very similar to that adopted by other
Bioconductor packages using the \Robject{expressionSet} and
\Robject{AnnotatedDataFrame} structures familiar to most Bioconductor
users.

\subsection{The \Rclass{flowFrame} Class}

The basic unit of manipulation in \Rpackage{flowCore} is the
\Rclass{flowFrame}, which corresponds roughly with a single
``FCS'' file exported from the flow cytometer's acquisition
software. At the moment we support FCS file versions 2.0
through 3.0, and we expect to support FCS4/ACS1 as soon as
the specification has been ratified.

\subsubsection{Data elements}

The primary elements of the \Robject{flowFrame} are the
\Robject{exprs} and \Robject{parameters} slots, which
contain the event-level information and column metadata
respectively. The event information, stored as a single
matrix, is accessed and manipulated via the
\Rfunction{exprs()} and \Rfunction{exprs<-} methods,
allowing \Robject{flowFrame}s to be stitched together if
necessary (for example, if the same tube has been collected
in two acquisition files for memory reasons).

The \Robject{parameters} slot is an
\Rclass{AnnotatedDataFrame} that contains information
derived from an FCS file's ``\$P<n>'' keywords, which
describe the detector and stain information. The entire list
is available via the \Rfunction{parameter()} method, but
more commonly this information is accessed through the
\Rfunction{names}, \Rfunction{featureNames} and
\Rfunction{colnames} methods. The \Rfunction{names} function
returns a concatenated version of \Rfunction{names} and
\Rfunction{featureNames} using a format similar to the one
employed by most flow cytometry analysis software. The
\Rfunction{colnames} method returns the detector names,
often named for the fluorochrome detected, while the
\Rfunction{featureNames} methods returns the description
field of the parameters, which will typically be an
identifier for the antibody.

The \Rfunction{keyword} method allows access to the raw FCS
keywords, which are a mix of standard entries such as
``SAMPLE ID,'' vendor specific keywords and user-defined
keywords that add more information about an experiment. In
the case of plate-based experiments, there are also one or
more keywords that identify the specific well on the plate.

Most vendor software also include some sort of unique
identifier for the file itself. The specialized methods
\Rfunction{identifier} attempts to locate an appropriate
globally unique identifier that can be used to uniquely
identify a frame. Failing that, this method will return the
original file name offering some assurance that this frame
is at least unique to a particular session.

\subsubsection{Reading a \Robject{flowFrame}}

FCS files are read into the R environment via the
\Rfunction{read.FCS} function using the standard connection
interface---allowing for the possibility of accessing FCS
files hosted on a remote resource as well as those that have
been compressed or even retrieved as a blob from a database
interface. FCS files (version 2.0 and 3.0) and LMD (List
Mode Data) extensions are currently supported.

There are also several immediate processing options
available in this function, the most important of which is
the \Rfunction{transformation} parameter, which can either
``linearize'' (the default) or ``scale'' our data. To
see how this works, first we will examine an FCS file
without any transformation at all:

<<ReadFiles, echo=true, results=verbatim>>=
file.name <- system.file("extdata","0877408774.B08", package="flowCore")
x <- read.FCS(file.name, transformation=FALSE)
summary(x)
@

As we can see, in this case the values from each parameter
seem to run from $0$ to $1023$ ($2^{10}-1$).  However,
inspection of the ``exponentiation'' keyword (\$P<n>E)
reveals that some of the parameters (3 and 4) have been
stored in the format of the form $a\times{10^{x/R}}$ where
$a$ is given by the first element of the string.

<<SearchKeywords, echo=true,results=verbatim>>=
keyword(x,c("$P1E", "$P2E", "$P3E", "$P4E"))
@

The default ``linearize'' transformation option will convert these to,
effectively, have a ``\$P<n>E'' of ``0,0'':

<<PrintSummary, echo=true, results=verbatim>>=
summary(read.FCS(file.name))
@

Finally, the ``scale'' option will both linearize values as
well as ensure that output values are contained in $[0,1]$,
which is the proposed method of data storage for the
ACS1.0/FCS4.0 specification:

<<PrintSummary2, echo=true,results=verbatim>>=
summary(read.FCS(file.name,transformation="scale")) 
@

Another parameter of interest is the \Rfunction{alter.names}
parameter, which will convert the parameter names into more
``R friendly'' equivalents, usually by replacing ``-'' with
``.'':

<<ReadFiles2, echo=true,results=verbatim>>=
read.FCS(file.name,alter.names=TRUE) 
@

When only a particular subset of parameters is desired the
\Rfunction{column.pattern} parameter allows for the specification of a
regular expression and only parameters that match the regular
expression will be included in the frame. For example, to include on
the Height parameters:

<<RedFiles3, echo=true, results=verbatim>>= 
x <- read.FCS(file.name, column.pattern="-H") 
x 
@

\textit{Note that \Rfunction{column.pattern} is applied after
  \Rfunction{alter.names} if it is used.}

Finally, only a sample of lines can be read in case you need a quick
overview of a large serie of files.
<<RedFiles3, echo=true, results=verbatim>>= 
lines <- sample(100:500, 50)
y <- read.FCS(file.name, which.lines = lines) 
y 
@

\subsubsection{Visualizing a \Rclass{flowFrame}}

Much of the more sophisticated visualization of \Rclass{flowFrame} and
\Rclass{flowSet} objects, including an interface to the
\Rpackage{lattice} graphics system is implemented by the
\Rpackage{flowViz} package, also included as part of
Bioconductor. However, \Rpackage{flowCore} does implement some basic
plotting facilities using the standard \Rfunction{plot} function. The
basic plot provides a simple pairs plot of all parameters:

<<Plot1, fig=true, echo=true>>=
plot(x)
@

To control the parameters being plotted we can supply a \Robject{y}
value in the form of a character vector. If we choose exactly two
parameters this will create a bivariate density plot.

<<Plot2, fig=true, echo=true>>=
plot(x,c("FL1-H", "FL2-H"))
@

However, if we only supply a single parameter we instead get
a univariate histogram, which also accepts the usual
histogram arguments:

<<plot3, fig=true, echo=true>>=
plot(x, "FL1-H", breaks=256)
@

\subsection{The \Rclass{flowSet} Class}

Most experiments consist of several \Rclass{flowFrame} objects, which
are organized using a \Rclass{flowSet} object. This class provides a
mechanism for efficiently hosting the \Rclass{flowFrame} objects with
minimal copying, reducing memory requirements, as well as ensuring
that experimental metadata stays properly to the appropriate
\Rclass{flowFrame} objects.

\subsubsection{Creating a \Rclass{flowSet}}

To facilitate the creation of \Rclass{flowSet} objects from a variety
of sources, we provide a means to coerce \Rclass{list} and
\Rclass{environment} objects to a \Rclass{flowSet} object using the
usual coercion mechanisms. For example, if we have a directory
containing FCS files we can read in a list of those files and create a
\Rclass{flowSet} out of them:

<<Frames, echo=true, results=verbatim>>=
frames <- lapply(dir(system.file("extdata", "compdata", "data",
                                 package="flowCore"), full.names=TRUE), 
                 read.FCS)
as(frames, "flowSet")
@

Note that the original list is unnamed and that the resulting sample
names are not particularly meaningful. If the list is named, the list
constructed is much more meaningful. One such approach is to employ
the \Rfunction{keyword} method for \Rclass{flowFrame} objects to
extract the ``SAMPLE ID'' keyword from each frame:

<<Frames, echo=true,results=verbatim>>=
names(frames) <- sapply(frames, keyword, "SAMPLE ID")
fs <- as(frames, "flowSet")
fs
@

\subsubsection{Working with experimental metadata}

Like most Bioconductor organizational classes, the \Rclass{flowSet}
has an associated \Rclass{AnnotatedDataFrame} that provides metadata
not contained within the \Rclass{flowFrame} objects themselves. This
data frame is accessed and modified via the usual
\Rfunction{phenoData} and \Rfunction{phenoData<-} methods. You can
also generally treat the phenotypic data as a normal data frame to add
new descriptive columns. For example, we might want to track the
original filename of the frames from above in the phenotypic data for
easier access:

<<metaData, echo=true, results=verbatim>>=
phenoData(fs)$Filename <- fsApply(fs,keyword, "$FIL")
pData(phenoData(fs))
@

Note that we have used the \Rclass{flowSet}-specific iterator,
\Rfunction{fsApply}, which acts much like \Rfunction{sapply} or
\Rfunction{lapply}. Additionally, we should also note that the
\Rfunction{phenoData} data frame \textbf{must} have row names that
correspond to the original names used to create the \Rclass{flowSet}.

\subsubsection{Bringing it all together: \Rfunction{read.flowSet}}

Much of the functionality described above has been packaged into the
\Rfunction{read.flowSet} convenience function. In it's simplest
incarnation, this function takes a \Rfunction{path}, that defaults to
the current working directory, and an optional \Rfunction{pattern}
argument that allows only a subset of files contained within the
working directory to be selected. For example, to read a
\Rclass{flowSet} of the files read in by \Robject{frame} above:

<<ReadFlowSet, echo=true,results=verbatim>>=
read.flowSet(path = system.file("extdata", "compdata", "data", 
             package="flowCore"))
@

\Rfunction{read.flowSet} will pass on additional arguments meant for
the underlying \Rfunction{read.FCS} function, such as
\Rfunction{alter.names} and \Rfunction{column.pattern}, but also
supports several other interesting arguments for conducting initial
processing:
\begin{description}
\item[files]{ An alternative to the \Rfunction{pattern} argument, you
    may also supply a vector of filenames to read.}
\item[name.keyword]{ Like the example in the previous section, you may
    specify a particular keyword to use in place of the filename when
    creating the \Rclass{flowSet}.}
\item[phenoData]{ If this is an \Rclass{AnnotatedDataFrame}, then this
    will be used in place of the data frame that is ordinarily
    created. Additionally, the row names of this object will be taken
    to be the filenames of the FCS files in the directory specified by
    \Robject{path}. This argument may also be a named list made up of
    a combination of \Robject{character} and \Robject{function}
    objects that specify a keyword to extract from the FCS file or a
    function to apply to each frame that will return a result. }
\end{description}

To recreate the \Rclass{flowSet} that we created by hand from the last
section we can use \Rfunction{read.flowSet}s advanced functionality:

<<ReadFowSet2, echo=true,results=verbatim>>=
fs <- read.flowSet(path=system.file("extdata", "compdata", "data",
                   package="flowCore"), name.keyword="SAMPLE ID",
                   phenoData=list(name="SAMPLE ID", Filename="$FIL"))
fs
pData(phenoData(fs))
@

\subsubsection{Manipulating a \Rclass{flowSet}}

You can extract a \Rclass{flowFrame} from a \Rclass{flowSet} object in
the usual way using the \Rfunction{[[} or \Rfunction{\$} extraction
  operators. On the other hand using he \Rfunction{[} extraction
  operator returns a new \Rclass{flowSet} by \textbf{copying} the
  environment. However, simply assigning the \Rclass{flowFrame} to a
  new variable will \textbf{not} copy the contained frames.

  The primary iterator method for a \Rclass{flowSet} is the
  \Rfunction{fsApply} method, which works more-or-less like
  \Rfunction{sapply} or \Rfunction{lapply} with two extra options. The
  first argument, \Robject{simplify}, which defaults to
  \Robject{TRUE}, instructs \Rfunction{fsApply} to attempt to simplify
  it's results much in the same way as \Rfunction{sapply}. The primary
  difference is that if all of the return values of the iterator are
  \Rclass{flowFrame} objects, \Rfunction{fsApply} will create a new
  \Rclass{flowSet} object to hold them. The second argument,
  \Robject{use.exprs}, which defaults to \Robject{FALSE} instructs
  \Rfunction{fsApply} to pass the expression matrix of each frame
  rather than the \Rclass{flowFrame} object itself. This allows
  functions to operate directly on the intensity information without
  first having to extract it.

  As an aid to this sort of operation we also introduce the
  \Rfunction{each\_row} and \Rfunction{each\_col} convenience
  functions that take the place of \Rfunction{apply} in the
  \Rfunction{fsApply} call. For example, if we wanted the median value
  of each parameter of each \Rclass{flowFrame} we might write:

<<fsApply1, echo=true, results=verbatim>>=
fsApply(fs, each_col, median)
@

which is equivalent to the less readable

<<fsApply2, echo=true,results=verbatim>>=
fsApply(fs,function(x) apply(x, 2, median), use.exprs=TRUE)
@

In this case, the \Robject{use.exprs} argument is not required in the
first case because \Rfunction{each\_col} and \Rfunction{each\_row} are
methods and have been defined to work on \Rclass{flowFrame} objects by
first extracting the intensity data.

\section{Transformation}

\Rpackage{flowCore} features two methods of transforming parameters
within a \Rclass{flowFrame}: inline and out-of-line. The inline
method, discussed in the next section has been developed primarily to
support filtering features and is strictly more limited than the
out-of-line transformation method, which uses R's
\Rfunction{transform} function to accomplish the filtering. Like the
normal \Rfunction{transform} function, the \Rclass{flowFrame}is
considered to be a data frame with columns named for parameters of the
FCS file. For example, if we wished to plot our first
\Rclass{flowFrame}'s first two fluorescence parameters on the log
scale we might write:

<<Transfo1, fig=true, echo=true>>=
plot(transform(fs[[1]], `FL1-H`=log(`FL1-H`), `FL2-H`=log(`FL2-H`)), 
     c("FL1-H","FL2-H"))
@

Like the usual \Rfunction{transform} function, we can also create new
parameters based on the old parameters, without destroying the old

<<Transfo2, fig=true, prefix=FALSE>>=
plot(transform(fs[[1]], log.FL1.H=log(`FL1-H`), 
               log.FL2.H=log(`FL2-H`)), c("log.FL1.H", "log.FL2.H"))
@

\subsection{Standard Transforms}

Though any function can be used as a transform in both the out-of-line
and inline transformation techniques, \Rpackage{flowCore} provides a
number of parameterized transform generators that correspond to the
transforms commonly found in flow cytometry and defined in the
Transformation Markup Language (Transformation-ML). Briefly, the
predefined transforms are:

\begin{description}
\item[truncateTransform]{ $y = \left\{\begin{array}{ll}
a & x < a \\
x & x \geq a \\
\end{array}\right.$ }
\item[scaleTransform]{$ f(x) = \frac{x-a}{b-a}$}
\item[linearTransform]{ $ f(x) = a+bx$ }
\item[quadraticTransform]{ $ f(x) = ax^2 + bx + c$ }
\item[lnTransform]{$ f(x) = \log\left(x\right)\frac{r}{d} $}
\item[logTransform]{$ f(x) = \log_b\left(x\right)\frac{r}{d} $}
\item[biexponentialTransform]{ $ f^{-1}(x) = ae^{bx}-ce^{dx}+f$}
\item[logicleTransform]{ A special form of the biexponential transform with parameters selected by the data.}
\item[arcsinhTransform]{ $ f(x) = asinh\left(a+bx\right)+c$}
\end{description}

To use a standard transform, first we create a transform function via the constructors supplied by \Rpackage{flowCore}:

<<Transfo3, echo=true>>=
aTrans <- truncateTransform("truncate at 1", a=1)
aTrans
@
which we can then use on the parameter of interest in the usual way
<<Transfo4, echo=true,results=verbatim>>=
transform(fs,`FL1-H`=aTrans(`FL1-H`))
@

\section{Filtering}

The most common task in the analysis of flow cytometry data is usual
some form of filtering operation, either to obtain summary statistics
about the number of events that meet a certain criteria or to perform
further analysis on a subset of the data.

\subsection{Standard Filters}

Most filtering operations are a composition of one or more common
filtering operations. Like transformations, \Rpackage{flowCore}
includes a number of built-in common flow cytometry filters.

The simplest of these filters are the geometric filters, which
correspond to those typically found in interactive flow cytometry
software:
\begin{description}
\item[rectangleGate]{Describes a cubic shape in one or more
    dimensions--a rectangle in one dimension is simply an interval
    gate.}
\item[polygonGate]{Describes an arbitrary two dimensional polygonal
    gate.}
\item[polytopeGate]{Describes a region that is the convex hull of the
    given points. This gate can exist in dimensions higher than $2$,
    unlike the \Robject{polygonGate}.}
\item[ellipsoidGate]{Describes an ellipsoidal region in two or more
    dimensions}
\end{description}

These gates are all described in more or less the same manner (see man
pages for more details):

<<rectGate, echo=true, results=verbatim>>=
rectGate <- rectangleGate(filterId="Fluorescence Region", 
                          "FL1-H"=c(50, 100), "FL2-H"=c(50, 100))
@

We also include some data-driven filters not usually found in flow
cytometry software:

\begin{description}
\item[norm2Filter]{ A robust method for finding a region that most
    resembles a bivariate Normal distribution. }
\item[kmeansFilter]{ Identifies populations based on a one dimensional
    k-means clustering operation. Allows the specification of
    \textbf{multiple} populations. }
\end{description}

\subsection{Count Statistics}

When we have constructed a filter, we can apply it in two basic
ways. The first is to collect simple summary statistics on the number
and proportion of events considered to be contained within the gate or
filter. This is done using the \Rfunction{filter} method. The first
step is to apply our filter to some data

<<echo=true,results=verbatim>>=
result = filter(fs[[1]],rectGate)
result
@

As we can see, we have returned a \Rclass{filterResult} object, which
is in turn a filter allowing for reuse in, for example, subsetting
operations. To obtain count and proportion statistics, we take the
summary of this \Rclass{filterResult}, which returns a list of summary
values:

<<Summary3, echo=true, results=verbatim>>=
summary(result)
summary(result)$n
summary(result)$true
summary(result)$p
@

A filter which contains multiple populations, such as the
\Rclass{kmeansFilter}, can return a list of summary lists:

<<SummarFilter, echo=true, results=verbatim>>=
summary(filter(fs[[1]], kmeansFilter("myKMeans","FSC-H"=c("Low", "Medium", "High"))))
@

A filter may also be applied to an entire \Rclass{flowSet}, in which
case it returns a list of \Rclass{filterResult} objects:

<<echo=true,results=verbatim>>=
filter(fs,rectGate)
@

\subsection{Subsetting}

To subset or split a \Rclass{flowFrame} or \Rclass{flowSet}, we use
the \Rfunction{Subset} and \Rfunction{split} methods respectively. The
first, \Rfunction{Subset}, behaves similarly to the standard R
\Rfunction{subset} function, which unfortunately could not used. For
example, recall from our initial plots of this data that the
morphology parameters, Forward Scatter and Side Scatter contain a
large more-or-less ellipse shaped population. If we wished to deal
only with that population, we might use \Rfunction{Subset} along with
a \Rclass{norm2Filter} object as follows:

<<Norm2Filter, echo=true, results=verbatim>>=
morphGate <- norm2Filter(filterId="MorphologyGate", "FSC-H", "SSC-H", scale=2)
smaller <- Subset(fs, morphGate)
fs[[1]]
smaller[[1]]
@

Notice how the \Robject{smaller} \Rclass{flowFrame} objects contain
fewer events. Now imagine we wanted to use a \Rclass{kmeansFilter} as
before to split our first fluorescence parameter into three
populations. To do this we employ the \Rfunction{split} function:

<<Split, echo=true, results=verbatim>>=
split(smaller[[1]], kmeansFilter("myKMeans","FSC-H"=c("Low","Medium","High")))
@

or for an entire \Rclass{flowSet}

<<Split2, echo=true, results=verbatim>>=
split(smaller, kmeansFilter("myKMeans", "FSC-H"=c("Low", "Medium", "High")))
@

\subsection{Combining Filters}

Of course, most filtering operations consist of more than one gate. To
combine gates and filters we use the standard R Boolean operators: \&,
| and ! to construct an intersection, union and complement
respectively:

<<CombineFilter, echo=true, results=verbatim>>=

rectGate & morphGate
rectGate | morphGate
!morphGate

@

we also introduce the notion of the subset operation, denoted by
either \%subset\% or \%\&\%. This combination of two gates first
performs a subsetting operation on the input \Rclass{flowFrame} using
the right-hand filter and then applies the left-hand filter. For
example,

<<Summary5, echo=true, results=verbatim>>=
summary(filter(smaller[[1]],rectGate %&% morphGate))
@

first calculates a subset based on the \Robject{morphGate} filter and
then applies the \Robject{rectGate}.

\subsection{Transformation Filters}

Finally, it is sometimes desirable to construct a filter with respect
to transformed parameters. To allow for this in our filtering
constructs we introduce a special form of the \Rfunction{transform}
method along with another filter combination operator \%on\%, which
can be applied to both filters and \Rclass{flowFrame} or
\Rclass{flowSet} objects. To specify our transform filter we must
first construct a transform list using a simplified version of the
\Rfunction{transform} function:

<<Transfo5, echo=true,results=verbatim>>=
tFilter <- transform("FL1-H"=log,"FL2-H"=log)
tFilter
@

Note that this version of the transform filter does not take
parameters on the right-hand side--the functions can only take a
single vector that is specified by the parameter on the left-hand
side. In this case those parameters are ``FL1-H'' and ``FL2-H.'' The
function also does not take a specific \Rclass{flowFrame} or
\Rclass{flowSet} allowing us to use this with any appropriate data. We
can then construct a filter with respect to the transform as follows:

<<TectGate3, echo=true, results=verbatim>>=
rect2 <- rectangleGate(filterId="Another Rect", "FL1-H"=c(1,2), "FL2-H"=c(2,3)) %on% tFilter
rect2
@

Additionally, we can use this construct directly on a
\Rclass{flowFrame} or \Rclass{flowSet} by moving the transform to the
left-hand side and placing the data on the right-hand side:

<<Plot6, fig=true,echo=true>>=
plot(tFilter %on% smaller[[1]],c("FL1-H","FL2-H"))
@

which has the same effect as the $\log$ transform used earlier.

\section{filterSet: Organizing Filtering Strategies}

\subsection{Building Filter Sets}

In addition to allowing for sequential filtering, \Rpackage{flowCore}
also provides a \Rclass{filterSet} object that serves as an analogue
to \Rclass{flowSets} for filters. The primary use of this object is to
organize and manipulate complex gating strategies. For example, recall
from the filtering section we had two gates, a ``morphologyGate'' and
a ``Fluorescence Region'' gate that we used to demonstrate the various
logical operators available for gates. To recreate this example as a
\Rclass{filterSet}, we would do the following:

<<FilterSet1, echo=true, results=verbatim>>=
fset1 = filterSet(
	rectangleGate(filterId="Fluorescence Region","FL1-H"=c(50,100),"FL2-H"=c(50,100)),
	norm2Filter(filterId="Morphology Gate","FSC-H","SSC-H",scale=2),
	~ `Fluorescence Region` & `Morphology Gate`,
	~ `Fluorescence Region` | `Morphology Gate`,
	Debris ~ ! `Morphology Gate`,
	~ `Fluorescence Region` %&% `Morphology Gate`
)
fset1

@

There are two features of note in \Rfunction{filterSet}, which can
also take a single list argument for programmatic creation of
\Rclass{filterSet} objects. The first is that there is a formula
interface for the creation of \Rclass{filter} objects from
operators. The formula interface can be used with or without a
left-hand side, which specifies an optional filter identifier. If the
filter identifier is not specified, as is the case with all but the
``Debris'' gate in the example above the usual default filter
identifier is constructed. Non-formula gates can also have an optional
name specified which overrides the filter identifier specified in the
gate constructor. This is discouraged at this time as it leads to
confusion when creating new filters.

\subsection{Manipulating Filter Sets}

Manipulating a \Rclass{filterSet} is done using the normal list
element replacement methods, though we do provide a special case where
replacing the ``""'' element or the ``NULL'' element causes the
\Rclass{filterSet} to use the ``filterId'' slot of the incoming filter
to choose a name. Similarly, specifying an identifier will override
any ``filterId'' slot in the original filter.

Additionally, there are several convenience functions and methods
available for manipulating a \Rclass{filterSet}. The \Rfunction{names}
function provides a list of filter identifiers in an arbitrary
order. The \Rfunction{sort} lists the filter identifiers according to
a topological sort of the filter dependencies such that parent filters
are always evaluated before their children. Filters without
dependencies are provided in an arbitrary relative ordering. To obtain
the adjacency matrix for this sorting, the option
``dependencies=TRUE'' supplied to the \Rfunction{sort} function will
attach an ``AdjM'' attribute to the resulting character vector.

\subsection{Using Filter Sets}

Though they are not explicitly filters, meaning that they cannot play
a role in the construction of subfilters via operators,
\Rclass{filterSet} objects do define \Rfunction{filter} and
\Rfunction{split} methods that allow for the two most common use
cases. Additionally, a filter can be selected from the
\Rclass{filterSet} using ``[['' to perform a \Rfunction{Subset}
operation if desired.

For standard filtering operations, the \Rclass{filterSet} can yield
better performance than a manual filtering operation because it
ensures that each parent filter is only evaluated once per call to
\Rfunction{filter} so that if many subfilters rely on a particular,
likely expensive, filtering operation they use the original result
rather than recalculating each time. The individual
\Rclass{filterResult} objects may be later extracted for further
analysis of desire, making the \Rclass{filterSet} the logical choice
for most filtering operations. Additionally, though not presently
implemented, it can allow visualization methods to make more
intelligent choices about display when rendering gating strategies
because the disposition of the entire filtering graph is known and the
filter results are already available.

Like any other filtering operation, \Rfunction{summary} methods are
also available to collect the usual sorts of count statistics for a
filtering operation:

<<FilterSet2, echo=TRUE, verbatim=TRUE>>=

f = filter(fs,fset1)
f
@

When splitting a \Rclass{flowFrame} using \Rfunction{split}, we have
two options. The first is to include all of the resultant subsets in
the usual way:

<<FilterSet3, echo=TRUE, verbatim=TRUE>>=

split(fs[[1]],fset1,flowSet=TRUE)

@

While this provides access to all subsets, for automated analysis we
will often only wish to work with the ``leaf'' gates (i.e. those
without children of their own). In this case, we can specify
``drop=TRUE'' to obtain only the leaf gates:

<<FilterSet4, echo=TRUE, verbatim=TRUE>>=

split(fs[[1]],fset1,drop=TRUE,flowSet=TRUE)

@

Note that in both cases, we use the ``flowSet=TRUE'' option that has
been added to the \Rfunction{split} function to automatically create a
\Rclass{flowSet} object from the resulting \Rclass{flowFrame}
objects. For \Rfunction{filterSet} objects, the \Rfunction{split}
function can take advantage of the added structure in this case to
provide some information about the filtering operation itself in the
\Rclass{flowSet}'s metadata as well.

\clearpage
\bibliographystyle{plainnat} 
\bibliography{cytoref}
\end{document}