### 2008-07-25 paper/manuscript.tex
Changes Nolwenn suggested
git-svn-id: file:///home/git/hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@32910 bc3139a8-67e5-0310-9ffc-ced21a209358

\documentclass[12pt]{article}
\usepackage{url}
\usepackage[authoryear,round]{natbib}
\usepackage{graphicx}

\title{Outline flowCore paper}

\author{Florian Hahne*\\
  Nolwenn LeMeur*\\
  Byron Ellis\\
  Ryan Brinkman\\
  Perry Haaland\\
  Errol Strain\\
  Deepayan Sarkar\\
  Josef Spidlen\\
  Robert Gentleman
 }

\begin{document}

\maketitle

\section*{Introduction}
\subsection*{Computational tools for the analysis of FC-HCS data}
During the last several years, automation technologies have been
developed that enable the use of flow cytometry high content screening
(FC-HCS) to generate large, complex datasets in both basic and
clinical research applications \citep{brinkman2007hcf}. A serious
bottleneck in the interpretation of existing studies and the
application of FC-HCS to even larger, more complex problems is that
data management and data analysis methods have not advanced
sufficiently far from the methods developed for traditional
applications of flow cytometry (FCM) to small-scale, tube-based
studies.

Some of the consequences of this lag in development are difficulties
in maintaining the integrity and documentation of extremely large
datasets, assessing measurement quality, developing validated assays,
controlling the accuracy of gating techniques, automating complex gating
strategies, and aggregating statistical results across large study
sets for further analysis.  Among the most serious consequences of the
current situation is, however, that it is very difficult for new
analysis approaches to find there way into standard practice. We
believe that this barrier to the development and dissemination of new
analysis methods is one of the fundamental restraints on the future
expansion of FC-HCS in both clinical and research applications.

In this paper, we describe a set of flexible and well structured
computational tools to efficiently analyze FC-HCS. This set of tools
is based on the open source statistical software R \citep{Rmain}, in
conjunction with the Bioconductor Project \citep{BIOC}. Our intent is
to provide a shared research platform that enables bioinformaticians,
computer scientists, and statisticians to work collaboratively with
biologists and clinicians to develop novel methods
\citep{lizard2007fca} for FCM data analysis. A critical component of
this approach is the development and implementation of standards that
will facilitate the adoption of these methods by both the larger
research community and commercial software vendors.  Consequently, we
are optimistic that this software platform will play a critical role
in addressing the bottlenecks described above that limit further
advances in the application of FC-HCS.

The computational tools we have developed are distributed in the R
software language as the Bioconductor package flowCore. flowCore is a
freely available, highly functional and extensible FCM data analysis
platform that enables researchers to efficiently handle FC-HCS data
and encourages open development of tools for their coherent
analysis. Among the issues addressed by flowCore are computationally
efficient data structures and a range of specialized methods for
compensation, transformation, and gating.  flowCore runs on Windows,
Mac OS X and Linux/Unix operating systems. Conceptually, the
underlying data structures that build the core of our software can be
implemented in any programming language, and we chose the R language
for its obvious advantages in statistical programming. In this paper,
we describe the functionality and methods underlying flowCore, and
provide examples of their use.

\subsection*{R and Bioconductor}
R is a robust statistical programing environment which offers a wide
range of statistical and visualization methods developed for various
fields of application. In particular, the Bioconductor project aims to
provide R software modules for biological and clinical data analysis.
It is widely used to process such data, spanning a diversity of
research fields from genomics to proteomics to cell biology
\citep{gentleman2006bos}. The R language and Bioconductor are
particularly geared towards the analysis of large complex datasets
that typically arise from high throughput experiments such as
microarray gene expression analysis or FC-HCS. Using flow cytometry as
a high-throughput technology requires suitable software infrastructure
to facilitate the rapid handling of the data. We rely on two important
lessons learned from the field of gene expression data analysis: the
first being the importance of data structures that reflect the
underlying data and facilitate the manipulations that are of most
interest, while the second lesson being the importance of a modular
architecture that allows for many developers to extend and use the
underlying infrastructure and to combine its tools in complex
work flows.

\subsection*{Existing data standards and conventions}
Currently, data from FCM experiments are stored in single files
according to the Flow Cytometry Standard (FCS) \citep{seamer1997pnd}.
However, recent developments in high-throughput FCM are shifting the
focus of interest away from single-tube based measurements towards
large and complex experimental designs with dozens of covariates and
influencing factors. Self-contained and self-reflective data
structures are a prerequisite to allow for coordinated analysis of
such experiments. Many of the currently available software solutions
offer only limited support for such structures, or make use of
containers like XML or binary storage formats that are designed
specifically for particular user interfaces and hence are not easily
amenable to programmatical access. In addition, the closed-source
nature of these products makes it impractical to integrate them into
analysis pipelines.

Traditionally, the majority of FCM experiments have been analyzed by
manual data inspection in one or two dimensions, or by very basic
comparisons of summary statistics. We believe that these approaches,
in addition to being extremely expensive and labor intensive, do not
fully address the highly complex nature of FCM data; in particular,
they discard many of the fundamental aspects of the data, such as its
underlying distribution and its high-dimensional nature. Furthermore,
the subjective character of manual analyses are a major obstacle to
reproducibility. For FC-HCS data, unassisted manual inspection is
extremely time consuming, and robust statistical methods need to be
developed to point investigators to interesting aspects of the data,
or to potential problems. While the expert knowledge of immunologists
and researchers remains crucial for the understanding of FCM data, we
believe that collaboration with other research fields such as
statistics and computer science can greatly improve the relevance of
FCM in today's high-throughput paradigm.

These issues are currently being addressed by emerging new flow
cytometry data standards developed in collaboration with the ISAC Data
Standards Task Force.  For example, the Gating-ML Candidate
Recommendation (CR) \textit{Fixme Ryan or Joseph: How do we cite
  this?}  represents a proposal on how to form unambiguous XML-based
gate definitions, which can facilitate the interchange and validation
of data between different software packages. Gates may be ordered into
a hierarchical structure to describe a gating strategy. This enables
the encoding of a simple analytical work flow, in a way that allows us
to reconstruct the analysis programmatically.

The flowCore framework presented here offers import functionality for
raw data FCS files along with their complete set of file-specific meta
information (Figure~\ref{fig1:FrameWork}). Moreover, it is a software
implementation of the Gating-ML CR, which opens the possibility to
integrate flowCore in existing work flows and to communicate with every
other FCM tool that adheres to the proposed standard. Adherence to
standards also plays a critical role in the ability of new methods
based on flowCore to find there way back into the standard practices
for FCM data analysis. Note that the data structures defined in
flowCore aim to model large sets of FCM data in a natural and coherent
way; they are not specific to the R language and can be implemented in
any other programming language. Ideally, different software programs
would communicate on the level of these data structures rather than a
high-level markup language.

flowCore is not a GUI-driven software, and all operations are done
using a command line interface. Conceptually, it is possible to add a
more elaborate user interface on top of that, however the focus is on
a programmatic approach to enable the convenient development of novel
analysis methods and automation of complex analysis approaches. By
taking the burden of data management from the programmer, and by
providing well-defined programming interfaces (APIs) and a structured
class hierarchy, it is possible to readily test new ideas and to
easily extend the framework's functionality.

\section*{Materials and Methods}
\subsection*{Basic data structures}
\subsubsection*{flowFrame}
FlowCore's primary task is the representation and basic manipulation
of FCM data. This is accomplished through a data model very close to
that adopted by other Bioconductor packages. All information from a
single FCS file, i.e., the collection of events and the accompanying
meta data, is stored in one single container. We call the structure
that hold this data a flowFrame. Raw data values as well as associated
meta data of a flowFrame can be accessed programmatically. Most
commonly, these are descriptors of the stains used in the experiment
and the respective measurement channels, information about
compensation performed at the instrument side and any additional
keywords the user deems to be important to annotate the data. A number
of quality checks are performed during the creation of a flowFrame to
ensure data integrity.

\subsubsection*{flowSet}
In high-throughput FCM, many of the analysis tasks need to be
performed consistently across multiple flow measurements, hence we
introduce the concept of a collection of flowFrames called a
flowSet. One can envision a flowSet to be a container for multiple
flowFrames along with relevant information associated with each
individual frame such as descriptions of the cell sample or the
treatment to which the sample was subjected. FlowSets manage the
consistent application of operations on the individual flowFrames and
shift the burden of keeping score of the annotation from the user to
the infrastructure, thus reducing the risk of errors (\textit{e.g.},
mixups of sample labels). The flowSet structure can be readily
extended to incorporate the potentially complex annotation associated
with even larger FC-HCS experiments. Access to flowCore data
structures and methods has already enabled the development of new
Bioconductor packages, such as plateCore, that facilitate the analysis
of plate-based FC-HCS data and make handling the annotation and
processing the data easier.

%%Figure1 Diagram of the data structures and the basic operations
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{Figure1-flowCoreFrameWork.jpg}
\caption{\label{fig1:FrameWork}{For each experiment, FCS files (as
    individual flowFrames), phenotypic and meta data are stored in a
    flowSet. Each flowFrame in a flowSet corresponds to one FCS
    file. All basic operations (e.g., compensation, transformation,
    gating) can be applied to either single flowFrames or a flowSet
    simultaneously.}}
\end{figure}

\subsection*{Standard flow operations}
The basic operations in FCM analyses are typically the same: the data
need to be compensated (if not already done on the instrument side)
and transformed, and sup-populations of interest need to be selected
based on a set of (predominantly sequential) gates. All software
solutions for FCM analysis offer more or less elaborate support for
these kinds of operation in an interactive, GUI-driven interface. The
approach we have taken in flowCore is to abstractly describe these
operations and build a set of tools to perform them on both flowFrames
and flowSets. While transformation, and to a certain extent
compensation, are fairly routine operations with only limited
potential for improvement, being able to implement new methodologies
for gating of FCM data, and extending the capabilities of flowCore
through object oriented programming are features that clearly sets our
framework apart from other FCM analysis tools. By factoring out as
much of the bookkeeping as possible, programmers can focus on the
actual operations rather then having to deal with the tedious details
of data integration and access. Third-party methods can act on their
own as first-class citizens in the analysis framework, without
breaking the work flow or the basic infrastructure. This design allows
for the straightforward extension of flowCore's capabilities, and has
already fostered the development of a number of valuable add-ons
\citep{lo2008agf,sarkar2008ufv} \textit{(Fixme Errol: Will plateCore
  be in Bioconductor soon. It should be if we want to cite it.)}.

\subsubsection*{Transformation and compensation}
Data transformation is essential for both data visualization and
modelling \citep{lo2008agf}. The major transformations that are
routinely used in FCM analysis have been implemented in flowCore
(\textit{e.g.,} quadratic, log or arcsinh, see Table~\ref{table1} for a
complete list). Furthermore, the design of the R language makes it
easy to define arbitrary functions to apply to the data of individual
flowFrames or entire flowSets, respectively. Compensation is
available for both flowFrames and flowSets, and the software also
offers functionality to compute spillover or compensation matrices
from a set of appropriate compensation samples.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\multicolumn{2}{|c|}{Data Transformations} \\
\hline
linear & $ax + b$ \\
quadratic & $ax^2 + bx + c$ \\
natural logarithm & $ln(x)(r/d)$ \\
logarithm & $log(x)(r/d)$ \\
biexponential & need equation \\
logicle& need equation \\
truncate & $x_{x<=a} = a$ \\
scale & $(x-a)/(b-a)$ \\
split-scale & combination of linear and logarithm \\
arcsinh & $asinh(a + bx)+c$ \\
\hline
\end{tabular}
\caption{\label{table1}Data transformations implemented in flowCore.}
\end{center}
\end{table}

\subsubsection*{Gating}
In flowCore, gating operations are represented by classes that can be
extended in an object-oriented manner (Table~\ref{table2}). Basic gate
types such as rectangular gates, ellipses and polygon gates are
implemented as part of the framework. In addition, we introduce the
notion of data-driven gates, or filters, for which the necessary
parameters are computed based on the properties of the underlying
data, for instance by modelling data distribution or by density
estimation. The ability to programmatically access gates is a
prerequisite for automated or semi-automated gating. By utilizing a
unified interface for all different types of gates, the user is able
to subset data sets as well as to create summary statistics, for
instance the proportions of events falling in single gates or in the
combinations of gates. Complex combinations and hierarchies of gates
can be captured in objects of class filterSet, allowing to apply
multi-step gating strategies. The definition of gates in flowCore
follows the Gating-ML CR, thus any flowCore gating strategy can be
reproduced by any other software that also adheres to the standard.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\multicolumn{2}{|c|}{Gates} \\
\hline
rectangleGate & n-dimensional rectangular regions \\
quadGate & quadrant regions in two dimensions \\
polygonGate & polygonal regions in two dimensions \\
polytopeGate & generalization of polygon in n dimensions \\
ellipsoidGate & n-dimensional ellipsoid region \\
\hline
\multicolumn{2}{|c|}{Filters} \\
\hline
sampleFilter & random sub-sampling\\
expressionFilter & results of a boolean expression \\
kmeansFilter & K-means clustering \\
norm2Filter & bivariate normal distribution \\
curv1Filter & local density regions in 1D \\
curv2Filter & density regions in 2D \\
timeFilter & abnormal data acquisition over time \\
\hline
filterSet & gating strategies \\
\hline
\end{tabular}
\caption{\label{table2}Filter and gate classes implemented in flowCore.}
\end{center}
\end{table}

\subsection*{Related flow packages}
In addition to the flowCore package that offers basic infrastructure,
we have implemented a range of additional Bioconductor packages that
are dedicated to more specific tasks of FCM data analysis. The flowViz
package \citep{sarkar2008ufv} provides sophisticated data
visualization tools, that make use of multivariate trellis
plotting. These functions can be used to quickly generate customized
plots for extended cytometry data sets for both direct data inspection
and quality control. The objects annotation information can be used to
arrange the layout and composition of the plots. Furthermore, the
design and the API of the visualization software is very generic, and
users can readily extend its capabilities by providing self-defined
plotting functions.  The flowQ package offers more advanced quality
assurance (QA) methodology and a framework to create interactive
web-based reports of QA results. Most of flowUtil's content deals with
data import and output from and to flowCore and the flow-cytometry
specific standard markup languages. Finally, flowStats provides more
elaborate statistical methods that are relevant in the context of flow
cytometry data analysis.

\section*{Results}
The flowCore package has been successfully applied in the analysis of
several datasets \citep{gasparetto2004ice,brinkman2007hcf}, and
several other Bioconductor packages are making use of its code
base. As an example, \cite{lo2008agf} have recently developed an
automatic gating approach via robust model-based clustering using
flowCore's data model and infrastructure and implemented this in the
Bioconductor package flowClust. Another package, plateCore, providing
more specialized support for experiments conducted on microtitre
plates is under development.

A complete documentation of the flowCore software is way beyond the
scope of this publication. Much more comprehensive documentation and
users guide information with programmatic examples is available online
(\url{http://bioconductor.org/packages/2.2/bioc/html/flowCore.html})
and as part of the package distribution. Here, we only want to briefly
exemplify some of the software's key features, that is, the coherent
treatment of all samples in a potentially large experiment, the
concept of data-driven automated gating, the integration of existing
software into the framework, and the generation of publication-quality
graphics for data visualization.

Data analysis for most experiments usually begins with a quality
assurance step, and we can use functionality from the flowQ package to
create an HTML report that highlights potential quality
issues. Assuming that the data has already been imported as the
flowSet object ``dat'', the following simple line of code produces the
quite complex output:

\begin{verbatim}
> library(flowCore)
> library(flowQ)
> qaReport(dat, c("qaProcess.timeline", "qaProcess.timeflow", 
                  "qaProcess.cellnumber"))
\end{verbatim}

The report is interactive and provides drill-down to more detailed
aspects of the analysis, starting from a concise overview
(Figure~\ref{flowQ}). The design of flowCore's data model allows for a
coherent treatment of all the samples, hence we are able to compare
features between individuals, or between groups of individuals, based
on the available annotation information.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75\textwidth]{flowQ.jpg}
\caption{\label{flowQ}% 
  HTML quality assessment report generated by
  the flowQ package. Rows correspond to samples, columns to different
  quality checks.}
\end{figure}

Static gating for all samples in a high-throughput FCM experiment is
often not an option, since the measured variables tend to vary between
different treatments, over time or between different experiment
batches. Automated or data-driven gating tries to estimate the gating
regions from the underlying data, thus providing a fast objective
solution to the analysis of potentially very large and diverse data
sets \cite{lo2008agf}. One of the automated gating methods implemented
in flowCore is based on identifying areas of significant curvature in
a kernel density estimate of the data \citep{wand2008}. Assuming that
the regions of interest are of high density, the software is able to
reliably detect them in a one or two-dimensional density landscape.

\begin{verbatim}
> cf <- curv2Filter("FL1-H", "FL3-H")
> fres <- filter(dat, cf)
\end{verbatim}

Kernel density estimation is a well-known problem in statistical
computing, and a lot of effort has been invested in the development of
good software to address it. The modular design of flowCore allows to
easily integrate these existing solutions into our framework. In this
example, we directly use R code from the feature package developed by
\cite{wand2008}. Instead of re-writing existing code, we are able to
include it via the well tested distribution mechanism provided by R's
software package system. This process is bi-directional, and all
functionality implemented in flowCore is available to other package
authors, as we have seen with the afore-mentioned flowClust package.

We can chose one of the many visualization option from the flowViz
package to plot the results of the recent filtering operation. A very
basic matrix of density plots is shown in Figure \ref{xyplot}, where
each panel in the matrix represents the data of one individual
patient.

\begin{verbatim}
> xyplot(`FL1-H` ~ `FL3-H` | SampleID, data=dat, filter=fres)
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75\textwidth]{xyplot}
\caption{\label{xyplot}%
Scatterplot matrix of a single flowSet with outlines of the gating
regions identified by an automated gating operation.}
\end{figure}

\section*{Discussion}

Through flowCore, we have provided the FCM community a freely
available, highly functional, development and analysis platform for
high throughput data analysis. Our intent is to provide a platform for
the collaborative development of new analysis methods that will
facilitate the transition of these new methods to the larger flow
community. Our experience has been that the collaborative effort we
enable for devising new methodology has been proven beneficial for a
number of different biological and computational biology challenges.
We hope that our framework will be the foundation for fruitful shared
research by many collaborators from multiple scientific fields and
will help resolve bottlenecks that currently prevent the further
development of and deployment of FC-HCS to increasingly complex and
important scientific and clinical applications.

\section*{Acknowledgements}
This work was supported by NIH grant EB005034. RRB is a Michael Smith
Foundation for Health Research Scholar and a International Society for
Analytical Cytology Scholar.

\bibliographystyle{plainnat}  
\bibliography{flowCoreRef} 
\end{document}