### 2008-08-15 paper/manuscript.tex
changes suggested by Ryan and Josef
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@33253 bc3139a8-67e5-0310-9ffc-ced21a209358

\documentclass[12pt]{article}
\usepackage{url}
\usepackage[authoryear,round]{natbib}
\usepackage{graphicx} 
\newcommand{\Rpackage}[1]{{\textsf{#1}}}
\newcommand{\Rclass}[1]{{\textit{#1}}}

\title{flowCore: A Bioconductor software package for high throughput flow cytometry analysis}

\author{Florian Hahne*\\
  Nolwenn LeMeur*\\
  Byron Ellis\\
  Ryan R. Brinkman\\
  Perry Haaland\\
  Errol Strain\\
  Deepayan Sarkar\\
  Josef Spidlen\\
  Robert Gentleman
 }

\begin{document}

\maketitle

\section*{Abstract}
\subsubsection*{Background}
Recent advances in automation technologies have enabled the use of
flow-cytometry high content screening (FH-HCS). However, data
management and data analysis methods have not advanced sufficiently
far from the inital small-scale studies to support the large complex
data sets generated in FH-HCS applications.
\subsubsection*{Methods}
In this manuscript, we describe a set of flexible
computational tools for FC-HCS data analysis written 
in R and freely available from the Bioconductor Project.

\subsubsection*{Results}
The \Rpackage{flowCore} package provides the flow cytometry community
with a shared research platform that enables bioinformaticians,
computer scientists, and statisticians to work collaboratively with
biologists and clinicians to develop novel methods for FCM data
analysis.
\subsubsection*{Conclusions}
We hope that our framework will be the foundation for fruitful 
collaborative research involving multiple scientific fields and
that it will help facilitate the
development and deployment of FC-HCS methods to complex 
scientific and clinical applications.

\section*{Introduction}
\subsection*{Computational tools for the analysis of FC-HCS data}
During the last several years, automation technologies have been
developed that enable the use of FCM high content screening
(FC-HCS) to generate large, complex datasets in both basic and
clinical research applications \citep{brinkman2007hcf}. A serious
bottleneck in the interpretation of existing studies and the
application of FC-HCS to even larger, more complex problems is that
data management and data analysis methods have not advanced
sufficiently far from the methods developed for applications of flow
cytometry (FCM) to small-scale, tube-based studies.

Some of the consequences of this lag in development are difficulties
in maintaining the integrity and documentation of extremely large
datasets, assessing measurement quality, developing validated assays,
controlling the accuracy of gating techniques, automating complex
gating strategies, and aggregating statistical results across large
study sets for further analysis.  Among the most serious consequences
of the current situation is, however, that it is very difficult for
new analysis approaches to find their way into standard practice. We
believe that this barrier to the development and dissemination of new
analysis methods is one of the fundamental restraints on the future
expansion of FC-HCS in both clinical and research applications.

In this paper, we describe a set of flexible and well structured
computational tools to efficiently analyze FC-HCS. This set of tools
is based on the open source statistical software R \citep{Rmain}, in
conjunction with the Bioconductor Project \citep{BIOC}. Our intent is
to provide a shared research platform that enables bioinformaticians,
computer scientists, and statisticians to work collaboratively with
biologists and clinicians to develop novel methods for FCM data
analysis, a process deemed crucial by many for the further development
of the technology \citep{lizard2007fca}. A critical component of this
approach is the development and implementation of standards that will
facilitate the adoption of these methods by both the larger research
community and commercial software vendors.  Consequently, we are
optimistic that this software platform will play a critical role in
addressing the bottlenecks described above that limit further advances
in the application of FC-HCS.

The computational tools we have developed are distributed in the R
software language as the Bioconductor package
\Rpackage{flowCore}. \Rpackage{flowCore} is a freely available, highly
functional and extensible FCM data analysis platform that enables
researchers to efficiently handle FC-HCS data and encourages open
development of tools for their coherent analysis. \Rpackage{flowCore}
implements computationally efficient data structures and a range of
specialized methods addressing all components of a typical FCM
analysis work flow, including compensation, transformation, and
gating. \Rpackage{flowCore} runs on Windows, Mac OS X and Linux/Unix
operating systems. Conceptually, the underlying data structures that
build the core of our software can be implemented in any programming
language. We chose the R language for its advantages in statistical
programming and visualization. In this paper, we describe the
functionality and methods underlying \Rpackage{flowCore}, and provide
examples of their use.

\subsection*{R and Bioconductor}
R is a robust statistical programing environment which offers a wide
range of statistical and visualization methods developed for various
fields of application. In particular, the Bioconductor project
provides R software modules for biological and clinical data analysis.
It is widely used to process such data, spanning a diversity of
research fields including genomics, proteomics and cell biology
\citep{BIOC}. The R language and Bioconductor are particularly geared
towards the analysis of large complex datasets that typically arise
from high throughput experiments such as microarray gene expression
analysis or FC-HCS. Using FCM as a high-throughput technology requires
suitable software infrastructure to facilitate the rapid handling of
the data. In our implementation of \Rpackage{flowCore} we rely on two
important lessons learned from the field of gene expression data
analysis: the first being the importance of data structures that
reflect the underlying data and facilitate the manipulations that are
of most interest, while the second is the importance of a modular
architecture that allows for many developers to extend and use the
underlying infrastructure and to combine tools in complex work flows.

\subsection*{Existing data standards and conventions}
Traditionally, the majority of FCM experiments have been analyzed by
manual data inspection in one or two dimensions, or by very basic
comparisons of summary statistics. Most of the currently available
analysis tools are designed to reflect this workflow.  We believe that
these approaches, in addition to being expensive and labor intensive,
do not fully address the highly complex nature of FCM data; in
particular, they disregard many of the fundamental aspects of the
data, such as its underlying distribution and its high-dimensional
nature. Furthermore, the subjective character of manual analyses are a
major obstacle to reproducibility. In a recent study of flow
cytometric standardization involving 15 institutions, the mean
inter-laboratory coefficient of variation ranged from 17 to 44\%, even
though preparation was standardized and performed using the same
samples and reagents at each site \citep{Maecker2005}. For FC-HCS
data, unassisted manual inspection is extremely time consuming, and
robust statistical methods need to be developed to point investigators
to interesting aspects of the data, or to potential problems. While
the expert knowledge of immunologists and researchers remains crucial
for the understanding of FCM data, we believe that collaboration with
other research fields such as statistics and computer science can
greatly improve the relevance of FCM in today's high-throughput
paradigm.

Currently, data from FCM experiments are stored in single files
according to the Flow Cytometry Standard (FCS) \citep{seamer1997pnd}.
However, recent developments in high-throughput FCM are shifting the
focus of interest away from single-tube based measurements towards
large and complex experimental designs with dozens of covariates and
influencing factors. For example, experiments consist of large numbers
of samples from different patients, measured at different timepoints
\citep{brinkman2007hcf} or following different drug treatments
\citep{gasparetto2004ice}. Modern FCM data analysis tools have to deal
with an additional layer of sample metadata and they need to provide
infrastructure to process and to compare groups of samples in a
concise and coordinated manner.

The notion of classes from an object-oriented programming language
provides one coherent way to describe these richer data structures,
and functions or methods that work on them allow for interaction and
manipulation. Many of the currently available software solutions offer
only limited support for such self-contained structures, or make use
of containers like XML or binary storage formats that are designed
specifically for particular user interfaces and hence not easily
amenable to programmatical access. In addition, the closed-source
nature of these products often makes them impractical to integrate
into analysis pipelines. In this manuscript we describe classes for
FCM data analysis and their implementation in R, however, they could
just as easily be implemented in any other language (e.g., Java,
C++). Software written in those languages could use similar data
structures, thereby simplifying communication and the interchange of
data between analysis tools.

The \Rpackage{flowCore} framework presented here can import and
process raw data FCS files along with their complete set of
file-specific metadata (Figure~\ref{fig1:FrameWork}).  Moreover, it is
a software implementation of the Gating-ML Candidate Recommendation
(CR), an emerging ISAC standard developed in collaboration with the
ISAC Data Standards Task Force, which makes it possibile to integrate
\Rpackage{flowCore} in existing work flows and to communicate with any
other FCM tool that adheres to the proposed standard
\citep{SpidlenInPressCytometryA}. Adherence to standards also plays a
critical role in the ability of new methods based on
\Rpackage{flowCore} to find their way back into the standard practices
for FCM data analysis.

\Rpackage{flowCore} is not a GUI-driven software, and all operations
are done using a command line interface.  It is possible to add a more
elaborate user interface on top of this infrastructure, however the
focus in this paper is on a programmatic approach to enable the
convenient development of novel analysis methods and automation of
complex analysis approaches.  By taking the burden of data management
from the programmer, and by providing well-defined programming
interfaces (APIs), it is possible to readily test new ideas and to
easily extend the framework's functionality.

\section*{Materials and Methods}
\subsection*{Basic data structures}
\subsubsection*{flowFrame}

\Rpackage{flowCore}'s primary task is the representation and basic
manipulation of FCM data. This is accomplished through a data model
very close to that adopted by other Bioconductor packages. All
information from a single FCS file, i.e., the collection of events and
the accompanying metadata, is stored in one single container. We call
the structure that hold this data a \Rclass{flowFrame}. Raw data
values as well as associated metadata of a \Rclass{flowFrame} can be
accessed programmatically. Most commonly, the metadata consist of
descriptors of the stains used in the experiment and the respective
measurement channels, information about compensation performed at the
instrument side and any additional keywords the user deems to be
important to annotate the data. During the creation of a
\Rclass{flowFrame}, a number of quality checks are performed to ensure
data integrity.

\subsubsection*{flowSet}
In high-throughput FCM, many of the analysis tasks need to be
performed consistently across multiple samples, hence we introduce the
concept of a collection of \Rclass{flowFrames} called a
\Rclass{flowSet}. A \Rclass{flowSet} is a container for multiple
\Rclass{flowFrames} along with relevant information associated with
each individual frame such as descriptions of the cell sample, the
treatment to which the sample was subjected, or the location of that
sample in a 96 or 384 well plate.  \Rclass{flowSets} manage the
consistent application of operations on the individual
\Rclass{flowFrames} and shift the burden of keeping score of the
metadata from the user to the infrastructure, thus reducing the risk
of errors (\textit{e.g.}, mixups of sample labels). The
\Rclass{flowSet} structure can be readily extended to incorporate the
potentially complex metadata associated with even larger FC-HCS
experiments such as clinical trials, where hundreds of patients might
provide samples at different time points over the course of the
experiment.

%%Figure1 Diagram of the data structures and the basic operations
\begin{figure}
\centering
\includegraphics[width=0.9\textwidth]{Figure1-flowCoreFrameWork.jpg}
\caption{\label{fig1:FrameWork}{For each experiment, the content of
    the FCS files, phenotypic and metadata are stored in a
    \Rclass{flowSet}. Each \Rclass{flowFrame} in a \Rclass{flowSet}
    corresponds to one FCS file. All basic operations (e.g.,
    compensation, transformation, gating) can be applied to either
    single \Rclass{flowFrames} or a \Rclass{flowSet} simultaneously.}}
\end{figure}

\subsection*{Standard flow operations}

Typically, the basic operations in FCM analyses adhere to the
following common work flow: the data need to be compensated (if that
was not already done on the instrument) and transformed, and
sup-populations of interest need to be selected based on a set of
(predominantly sequential) gates. All software solutions for FCM
analysis offer support for these operations, typically in an
interactive, GUI-driven interface. In \Rpackage{flowCore} we have
taken the approach to abstractly describe these operations and build a
set of tools to perform them on both \Rclass{flowFrames} and
\Rclass{flowSets}. While transformation, and to a certain extent
compensation, are fairly routine operations with only limited
potential for improvement, being able to implement new methodologies
for gating of FCM data, and extending the capabilities of
\Rpackage{flowCore} through object oriented programming are features
that clearly sets our framework apart from other FCM analysis
tools. By factoring out as much of the bookkeeping as possible,
programmers can focus on the actual operations rather then having to
deal with the tedious details of data integration and
access. Third-party methods can act on their own as first-class
citizens in the analysis framework, without breaking the work flow or
the basic infrastructure. This design allows for the straightforward
extension of \Rpackage{flowCore}'s capabilities, and has already
fostered the development of a number of valuable add-ons
\citep{lo2008agf,sarkar2008ufv}.

\subsubsection*{Transformation and compensation}

Data transformation is essential for both data visualization and
modelling \citep{lo2008agf}. The major transformations that are
routinely used in FCM analysis have been implemented in
\Rpackage{flowCore} (\textit{e.g.,} log, bi-exponantioal or arcsinh,
see Table~\ref{table1} for a complete list). Furthermore, the design
of the R language makes it easy to define arbitrary functions to apply
to the data of individual \Rclass{flowFrames} or entire
\Rclass{flowSets}, respectively. Compensation is available for both
\Rclass{flowFrames} and \Rclass{flowSets}, and the software also
offers functionality to compute spillover or compensation matrices
from a set of appropriate single stain controls.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\multicolumn{2}{|c|}{Data Transformations} \\
\hline
linear & $ax + b$ \\
quadratic & $ax^2 + bx + c$ \\
natural logarithm & $ln(x)(r/d)$ \\
logarithm & $log(x)(r/d)$ \\
biexponential & $root(a*e^{(b*y)}-c*e^{(-d+y)}+f)$ \\
logicle& $root(Te^{-(m-w)}*(e^{(y-w)}-p^2e^{-(y-w)/p}+p^2-1)$ \\
truncate & $x_{x<=a} = a$ \\
scale & $(x-a)/(b-a)$ \\
split-scale & combination of linear and logarithm \\
arcsinh & $asinh(a + bx)+c$ \\
\hline
\end{tabular}
\caption{\label{table1}Data transformations implemented in
  \Rpackage{flowCore}. Within these formulas, $x$ is the variable
  corresponding to value being transformed, $y$ is the result of the
  transformation, $a$, $b$, $c$, $d$, $f$, $p$, $m$, $T$, and $w$, are
  constants affecting the transformation function, $e$ is the base of
  the natural logarithm, and $root$ is a root-finding algorithm.  }
\end{center}
\end{table}

\subsubsection*{Gating}
In \Rpackage{flowCore}, gating operations are represented by classes
that can be extended in an object-oriented manner
(Table~\ref{table2}). Basic gate types such as rectangular gates,
ellipses and polygon gates are implemented as part of the
framework. In addition, we introduce the notion of data-driven gates,
or filters, for which the necessary parameters are computed based on
the properties of the underlying data, for instance by modelling data
distribution or by density estimation. This approach is fundamentally
different from the traditional application of static gating regions
across samples, as it is able to take into accounts unforeseen changes
in signal intensities, such as drifts in the instrumentation
over time or sample variablity.

The ability to programmatically access gates is a prerequisite for
automated or semi-automated gating. By utilizing a unified interface
for all different types of gates, the user is able to subset data sets
as well as to create summary statistics, for example the proportions
of events falling in a single gate or in a combination of
gates. Complex combinations and hierarchies of gates can be captured
in objects of class \Rclass{filterSet}, allowing to apply multi-step
gating strategies. The definition of gates in \Rpackage{flowCore}
follows the Gating-ML CR, thus any \Rpackage{flowCore} gating strategy
can be reproduced by any other software that also adheres to the
standard.

Gating, as well as all other operations in \Rclass{flowCore}, can be
applied over each individual frame in a \Rclass{flowSet}, and summary
methods provide information about the outcome of these operations. In
addition, the result of a gating operation can be used to subset the
input \Rclass{flowFrame} or \Rclass{flowSet}, either by filtering out
negative events or by splitting in multiple subpopulations. This
design allows to easily combine all of \Rpackage{flowCore}'s
components into complex work flows.

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|l|}
\hline
\multicolumn{2}{|c|}{Gates} \\
\hline
rectangleGate & n-dimensional rectangular regions \\
quadGate & quadrant regions in two dimensions \\
polygonGate & polygonal regions in two dimensions \\
polytopeGate & generalization of polygon in n dimensions \\
ellipsoidGate & n-dimensional ellipsoid region \\
\hline
\multicolumn{2}{|c|}{Filters} \\
\hline
sampleFilter & random sub-sampling\\
expressionFilter & results of a boolean expression \\
kmeansFilter & K-means clustering \\
norm2Filter & bivariate normal distribution \\
curv1Filter & local density regions in 1D \\
curv2Filter & density regions in 2D \\
timeFilter & abnormal data acquisition over time \\
\hline
filterSet & gating strategies \\
\hline
\end{tabular}
\caption{\label{table2}Filter and gate classes implemented in
  \Rpackage{flowCore}.}
\end{center}
\end{table}

\subsection*{Related flow packages}

In addition to the \Rpackage{flowCore} package that offers basic
infrastructure, we have implemented a range of additional Bioconductor
packages that are dedicated to more specific tasks of FCM data
analysis. The \Rpackage{flowViz} package \citep{sarkar2008ufv}
provides sophisticated data visualization tools, that make use off
multivariate trellis plotting \citep{lattice}.  These functions can be
used to quickly generate customized plots for extended cytometry data
sets for both direct data inspection and quality control.  The objects
metadata information can be used to arrange the layout and composition
of the plots.  Furthermore, the design and the API of the
visualization software is very generic, and users can readily extend
its capabilities by providing self-defined plotting functions.  The
\Rpackage{flowQ} package offers more advanced quality assurance (QA)
methodology and a framework to create interactive web-based reports of
QA results. The \Rpackage{flowUtil} package implements data import and
export including flow-cytometry specific standard markup
language. Finally, the \Rpackage{flowStats} package provides elaborate
statistical methods that are relevant in the context of flow
cytometry data analysis. \cite{lo2008agf} have recently developed an
automatic gating approach via robust model-based clustering using
\Rpackage{flowCore}'s data model and infrastructure which is
implemented in the Bioconductor package \Rpackage{flowClust}. Another
package, \Rpackage{plateCore}, providing more specialized support for
experiments conducted on microtitre plates and facilitiating the
handling of spatial metadata, is under development.

\section*{Results}
The \Rpackage{flowCore} package has been successfully applied in the
analysis of several datasets
\citep{gasparetto2004ice,brinkman2007hcf}. A complete documentation of
the \Rpackage{flowCore} software is beyond the scope of this
publication.  Much more comprehensive documentation and users guide
information with programmatic examples is available online
(\url{http://bioconductor.org/packages/2.2/bioc/html/flowCore.html})
and as part of the package distribution. Here, we want to briefly
exemplify some of the software's key features, that is, the coherent
treatment of all samples in a potentially large experiment, the
concept of data-driven automated gating, the integration of existing
software into the framework, and the generation of publication-quality
graphics for data visualization.

Data analysis for most experiments usually begins with a quality
assurance step, and we can use functionality from the \Rpackage{flowQ}
package to create an HTML report that highlights potential quality
issues. Assuming that the data has already been imported as the
\Rclass{flowSet} object ``dat'', the following simple line of code
produces the output shown in Figure~\ref{flowQ}:

\begin{verbatim}
> library(flowCore)
> library(flowQ)
> qaReport(dat, c("qaProcess.timeline", "qaProcess.timeflow", 
                  "qaProcess.cellnumber"))
\end{verbatim}

The report is interactive and provides drill-down to more detailed
aspects of the analysis, starting from a concise overview. The design of \Rpackage{flowCore}'s data model
allows for a coherent treatment of all the samples, hence we are able
to compare features between individuals, or between groups of
individuals, based on the available metadata information.

\begin{figure}[htbp]
\centering
\includegraphics[width=0.98\textwidth]{flowQ.jpg}
\caption{\label{flowQ}%
  HTML quality assessment report generated by the flowQ package for a
  subset of data from an experiment focussing on Graft-Versus-Host
  Disease \citep{brinkman2007hcf}. Rows correspond to the samples in
  the set, columns to different quality checks.}
\end{figure}

Static gating for all samples in a high-throughput FCM experiment is
often impossible, since the measured variables tend to vary between
different treatments, over time or between different experiment
batches. Automated or data-driven gating has the potential to estimate
the gating regions from the underlying data, thus providing a fast
objective solution to the analysis of potentially very large and
diverse data sets \citep{lo2008agf}. One of the automated gating
methods implemented in \Rpackage{flowCore} is based on identifying
areas of significant curvature in a kernel density estimate of the
data \citep{wand2008}. Assuming that the regions of interest are of
high density, the software is able to reliably detect them in a one or
two-dimensional density landscape.

\begin{verbatim}
> cf <- curv2Filter("FL1-H", "FL3-H")
> fres <- filter(dat, cf)
\end{verbatim}

Kernel density estimation is a well-known problem in statistical
computing, and a lot of effort has been invested in the development of
good software to address it. The modular design of \Rpackage{flowCore}
allows to easily integrate these existing solutions into our
framework. In this example, we directly use R code from the
\Rpackage{feature} package developed by \cite{wand2008}. Instead of
re-writing existing code, we are able to include it via the well
tested distribution mechanism provided by R's software package
system. This process is bi-directional, and all functionality
implemented in \Rpackage{flowCore} is available to other package
authors, as we have seen with the afore-mentioned \Rpackage{flowClust}
package.

We can chose one of the many visualization option from the flowViz
package to plot the results of the recent filtering operation. A very
basic matrix of density plots is shown in Figure \ref{xyplot}, where
each panel in the matrix represents the data of one individual
patient.

\begin{verbatim}
> xyplot(`FL1-H` ~ `FL3-H` | SampleID, data=dat, filter=fres)
\end{verbatim}

\begin{figure}[htbp]
\centering
\includegraphics[width=0.75\textwidth]{xyplot}
\caption{\label{xyplot}%
  Scatterplot matrix of a single \Rclass{flowSet} from an experiement
  focussing on immune tolerance following kidney
  transplantation. Outlines of the gating regions identified by an
  automated gating operation are added on top of the density
  representation of the data.}
\end{figure}

\section*{Discussion}

Through \Rpackage{flowCore}, we have provided the FCM community an
open source, freely available, highly functional, standards compliant
development and analysis platform for high throughput data analysis.
Our intention is to provide a platform for the collaborative
development of new analysis methods that will facilitate the
transition of these new methods to the larger flow community.  Our
experience has been that the collaborative effort we enable for
devising new methodology has been proven beneficial for a number of
different biological and computational biology challenges.  We hope
that our framework will be the foundation for fruitful shared research
by many collaborators from multiple scientific fields and will help
resolve bottlenecks that currently prevent further development and
deployment of FC-HCS to increasingly complex and important scientific
and clinical applications.

\section*{Acknowledgements}
This work was supported by NIH grant EB005034 and by the Michael Smith
Foundation for Health Research. RRB is an ISAC Scholar.

\bibliographystyle{plainnat}  
\bibliography{flowCoreRef} 
\end{document}