### 2008-11-25 paper/bmc_article.tex
work on the new layout
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@35475 bc3139a8-67e5-0310-9ffc-ced21a209358

%% BioMed_Central_Tex_Template_v1.05
%%                                      %
%  bmc_article.tex            ver: 1.05 %
%                                       %

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%         <27 January 2006>           %%
%%                                     %%
%%                                     %%
%% Uses:                               %%
%% cite.sty, url.sty, bmc_article.cls  %%
%% ifthen.sty. multicol.sty		       %%
%%									   %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%	
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.pdf and the instructions for    %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\documentclass[10pt]{bmc_article}

% Load packages
\usepackage{cite} % Make references as [1-4], not [1,2,3,4]
\usepackage{url}  % Formatting web addresses  
\usepackage{ifthen}  % Conditional 
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
\urlstyle{rm}

\newcommand{\Rpackage}[1]{{\textsf{#1}}}
\newcommand{\Rclass}[1]{{\textit{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%   
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %% 
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %% 
%%  submitted article.                         %% 
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

\setlength{\topmargin}{0.0cm}
\setlength{\textheight}{21.5cm}
\setlength{\oddsidemargin}{0cm} 
\setlength{\textwidth}{16.5cm}
\setlength{\columnsep}{0.6cm}

\newboolean{publ}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                              %%
%% You may change the following style settings  %%
%% Should you wish to format your article       %%
%% in a publication style for printing out and  %%
%% sharing with colleagues, but ensure that     %%
%% before submitting to BMC that the style is   %%
%% returned to the Review style setting.        %%
%%                                              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%Review style settings
\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}

%Publication style settings
%\newenvironment{bmcformat}{\fussy\setboolean{publ}{true}}{\fussy}

% Begin ...
\begin{document}
\begin{bmcformat}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{flowCore a Bioconductor package for high throughput flow cytometry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Ensure \and is entered between all but   %%
%% the last two authors. This will be       %%
%% replaced by a comma in the final article %%
%%                                          %%
%% Ensure there are no trailing spaces at   %% 
%% the ends of the lines                    %%     	
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author{Florian Hahne\correspondingauthor$^1$%
       \email{Florian Hahne\correspondingauthor - fhahne@fhcrc.org}%
      \and
         Nolwenn LeMeur\correspondingauthor$^{1,2}$%
         \email{Nolwenn LeMeur\correspondingauthor - nlemeur@irisa.fr}
       \and 
         Ryan R. Brinkman$^3$%
         \email{Ryan R. Brinkman - rbrinkman@bccrc.ca}%
        \and
          Byron Ellis$^4$%
          \email{Byron Ellis - byron.ellis@gmail.com}
        \and
          Perry Haaland$^5$%
          \email{Perry Haaland - Perry\_D\_Haaland@bd.com}
        \and
          Deepayan Sarkar$^1$%
          \email{Deepayan Sarkar - dsarkar@fhcrc.org}
        \and
         Josef Spidlen$^3$%
         \email{Josef Spidlen - jspidlen@bccrc.ca}
        \and
         Errol Strain$^5$%
         \email{Errol Strain - estrain@gmail.com}  
       and
          Robert Gentleman$^1$%   
          \email{Robert Gentleman - rgentlem@fhcrc.org}
        }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\address{%
  \iid(1)Life Sciences Department, Computational Biology Program,%
  Division of Public Health Sciences, Fred Hutchinson Cancer Research Center,%
  1100 Fairview Ave. N, M2-B876, PO Box 19024,  Seattle, Washington 98109-1024\\
  \iid(2)EA SeRAIC INSERM, IRISA - Symbiose, Campus Beaulieu, Universit\'{e} de Rennes I, % 
  35042 Rennes Cedex, France \\
  \iid(3)Terry Fox Laboratory, British Columbia Cancer Agency Research Centre,% 
  675 West 10th Avenue, Vancouver, BC V5Z 1L3  Canada, Canada\\
  \iid(4)AdBrite Inc\\
  \iid(5)BD Biosciences
}%
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% The Section headings here are those for  %%
%% a Research article submitted to a        %%
%% BMC-Series journal.                      %%  
%%                                          %%
%% If your article is not of this type,     %%
%% then refer to the Instructions for       %%
%% authors on http://www.biomedcentral.com  %%
%% and change the section headings          %%
%% accordingly.                             %%   
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstract}
        % Do not use inserted blank lines (ie \\) until main body of text.
  \paragraph*{Background:} Recent advances in automation technologies
  have enabled the use of flow-cytometry high content screening
  (FH-HCS). However, data management and data analysis methods have
  not advanced sufficiently far from the inital small-scale studies to
  support the large complex data sets generated in FH-HCS
  applications.

  \paragraph*{Results:} We developped a set of flexible computational
  tools for FC-HCS data analysis written in R and freely available
  from the Bioconductor Project. The \Rpackage{flowCore} package
  provides the flow cytometry community with a shared research
  platform that enables bioinformaticians, computer scientists, and
  statisticians to work collaboratively with biologists and clinicians
  to develop novel methods for flow cytometry data analysis.

  \paragraph*{Conclusions:} We hope that our framework will be the
  foundation for fruitful collaborative research involving multiple
  scientific fields and that it will help facilitate the development
  and deployment of FC-HCS methods to complex scientific and clinical
  applications.
\end{abstract}

\ifthenelse{\boolean{publ}}{\begin{multicols}{2}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% The Section headings here are those for  %%
%% a Research article submitted to a        %%
%% BMC-Series journal.                      %%  
%%                                          %%
%% If your article is not of this type,     %%
%% then refer to the instructions for       %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and change the section headings          %%
%% accordingly.                             %% 
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Automation technologies developed during the last several years have
enabled the use of flow cytometry high content screening (FC-HCS) to
generate large, complex datasets in both basic and clinical research
applications \cite{brinkman2007hcf}. A serious bottleneck in the
interpretation of existing studies and the application of FC-HCS to
even larger, more complex problems is that data management and data
analysis methods have not advanced sufficiently far from the methods
developed for applications of flow cytometry (FCM) to small-scale,
tube-based studies \cite{mahnke2007omi}.
%% add one of the last roederer's paper that "request" algo and
%% automation for FC-HCS data analysis.

Some of the consequences of this lag are difficulties in maintaining
the integrity and documentation of extremely large datasets, assessing
measurement quality, developing validated assays, controlling the
accuracy of gating techniques, automating complex gating strategies,
and aggregating statistical results across large study sets for
further analysis. The most serious consequence, however, is that new
analysis approaches face difficulty in finding their way into standard
practice. We believe that this barrier to the development and
dissemination of new analysis methods is one of the fundamental
restraints on the future expansion of FC-HCS in both clinical and
research applications.

Another critical component is the development and implementation of
standards that will facilitate the adoption of these methods by both
the larger research community and commercial software vendors.
Traditionally, the majority of FCM experiments have been analyzed by
manual data inspection in one or two dimensions, or by very basic
comparisons of summary statistics. Most of the currently available
analysis tools are designed to reflect this workflow.  We believe that
these approaches, in addition to being expensive and labor intensive,
do not fully address the highly complex nature of FCM data; in
particular, they disregard many of the fundamental aspects of the
data, such as its underlying distribution and its high-dimensional
nature. Furthermore, the subjective character of manual analyses are a
major obstacle to reproducibility. In a recent study of flow
cytometric standardization involving 15 institutions, the mean
inter-laboratory coefficient of variation ranged from 17 to 44\%, even
though preparation was standardized and performed using the same
samples and reagents at each site \cite{Maecker2005}. For FC-HCS data,
unassisted manual inspection is extremely time consuming, and robust
statistical methods need to be developed to point investigators to
interesting aspects of the data, or to potential problems. While the
expert knowledge of immunologists and researchers remains crucial for
the understanding of FCM data, we believe that collaboration with
other research fields such as statistics and computer science can
greatly improve the relevance of FCM in today's high-throughput
paradigm.

In this paper, we describe a set of flexible and well structured
computational tools to efficiently analyze FC-HCS data. This set of
tools is based on the open source statistical software R \cite{Rmain}
and distributed as the Bioconductor package \Rpackage{flowCore}
\cite{BIOC}. Our intent is to provide a shared research platform that
enables bioinformaticians, computer scientists, and statisticians to
work collaboratively with biologists and clinicians to develop novel
methods for FCM data analysis, a process deemed crucial by many for
the further development of the technology \cite{lizard2007fca}.

%% Perhaps in Conclusion
%% Consequently, we are optimistic that this
%% software platform will play a critical role in addressing the
%% bottlenecks described above that limit advances in the application of
%% FC-HCS.
%%

%%%%%%%%%%%%%%%%%%
\section*{Implementation}
%% The computational tools we have developed are distributed in the R
%% software language as the Bioconductor package
%% \Rpackage{flowCore}. 
The Bioconductor package \Rpackage{flowCore} is a freely available,
highly functional, and extensible FCM data analysis platform that
enables researchers to efficiently handle FC-HCS data and encourages
open development of tools for their coherent
analysis. \Rpackage{flowCore} implements computationally efficient
data structures and a range of specialized methods addressing all
components of a typical FCM analysis work flow, including
compensation, transformation, and gating. \Rpackage{flowCore} runs on
Windows, Mac OS X, and Linux/Unix operating systems. Conceptually, the
underlying data structures that build the core of our software can be
implemented in any programming language. We chose the R language for
its advantages in statistical programming and visualization. 
%% In this paper, we describe the functionality and methods underlying
%% \Rpackage{flowCore}, and provide examples of their use.

\subsection*{Existing data standards and conventions}
Currently, data from FCM experiments are stored in single files
according to the Flow Cytometry Standard (FCS) \cite{seamer1997pnd}.
However, recent developments in high-throughput FCM are shifting the
focus of interest away from single-tube based measurements towards
large and complex experimental designs with dozens of covariates and
influencing factors. For example, experiments consist of large numbers
of samples from different patients, measured at different timepoints
\cite{brinkman2007hcf} or following different drug treatments
\cite{gasparetto2004ice}. Modern FCM data analysis tools have to deal
with an additional layer of sample metadata and they need to provide
infrastructure to process and to compare groups of samples in a
concise and coordinated manner.

The notion of classes from an object-oriented programming language
provides one coherent way to describe these richer data structures,
and functions or methods that work on them allow for interaction and
manipulation. Many of the currently available software solutions offer
only limited support for such self-contained structures, or make use
of containers like XML or binary storage formats that are designed
specifically for particular user interfaces and hence not easily
amenable to programmatical access. In addition, the closed-source
nature of these products often makes them impractical to integrate
into analysis pipelines. In this manuscript we describe classes for
FCM data analysis and their implementation in R, however, they could
just as easily be implemented in any other language (e.g., Java,
C++). Software written in those languages could use similar data
structures, thereby simplifying communication and the interchange of
data between analysis tools.

The \Rpackage{flowCore} framework presented here can import and
process raw data FCS files along with their complete set of
file-specific metadata (Figure~\ref{fig1:FrameWork}).  Moreover, it is
a software implementation of the Gating-ML Candidate Recommendation
(CR), an emerging ISAC standard developed in collaboration with the
ISAC Data Standards Task Force, which makes it possibile to integrate
\Rpackage{flowCore} in existing work flows and to communicate with any
other FCM tool that adheres to the proposed standard
\cite{SpidlenInPressCytometryA}. Adherence to standards also plays a
critical role in the ability of new methods based on
\Rpackage{flowCore} to find their way back into the standard practices
for FCM data analysis.

\Rpackage{flowCore} is not a GUI-driven software, and all operations
are done using a command line interface.  It is possible to add a more
elaborate user interface on top of this infrastructure, however the
focus in this paper is on a programmatic approach to enable the
convenient development of novel analysis methods and automation of
complex analysis approaches.  By taking the burden of data management
from the programmer, and by providing well-defined programming
interfaces (APIs), it is possible to readily test new ideas and to
easily extend the framework's functionality.

\subsection*{flowCore basic data structures}
\subsubsection*{flowFrame}
\Rpackage{flowCore}'s primary task is the representation and basic
manipulation of FCM data. This is accomplished through a data model
very close to that adopted by other Bioconductor packages. All
information from a single FCS file, i.e., the collection of events and
the accompanying metadata, is stored in one single container. We call
the structure that hold this data a \Rclass{flowFrame}. Raw data
values as well as associated metadata of a \Rclass{flowFrame} can be
accessed programmatically. Most commonly, the metadata consist of
descriptors of the stains used in the experiment and the respective
measurement channels, information about compensation performed at the
instrument side and any additional keywords the user deems to be
important to annotate the data. During the creation of a
\Rclass{flowFrame}, a number of quality checks are performed to ensure
data integrity.

\subsubsection*{flowSet}
In high-throughput FCM, many of the analysis tasks need to be
performed consistently across multiple samples, hence we introduce the
concept of a collection of \Rclass{flowFrames} called a
\Rclass{flowSet}. A \Rclass{flowSet} is a container for multiple
\Rclass{flowFrames} along with relevant information associated with
each individual frame such as descriptions of the cell sample, the
treatment to which the sample was subjected, or the location of that
sample in a 96 or 384 well plate.  \Rclass{flowSets} manage the
consistent application of operations on the individual
\Rclass{flowFrames} and shift the burden of keeping score of the
metadata from the user to the infrastructure, thus reducing the risk
of errors (\textit{e.g.}, mixups of sample labels). The
\Rclass{flowSet} structure can be readily extended to incorporate the
potentially complex metadata associated with even larger FC-HCS
experiments such as clinical trials, where hundreds of patients might
provide samples at different time points over the course of the
experiment.

\subsection*{Standard flow operations}

Typically, the basic operations in FCM analyses adhere to the
following common work flow: the data need to be compensated (if that
was not already done on the instrument) and transformed, and
sup-populations of interest need to be selected based on a set of
(predominantly sequential) gates. All software solutions for FCM
analysis offer support for these operations, typically in an
interactive, GUI-driven interface. In \Rpackage{flowCore} we have
taken the approach to abstractly describe these operations and build a
set of tools to perform them on both \Rclass{flowFrames} and
\Rclass{flowSets}. While transformation, and to a certain extent
compensation, are fairly routine operations with only limited
potential for improvement, being able to implement new methodologies
for gating of FCM data, and extending the capabilities of
\Rpackage{flowCore} through object oriented programming are features
that clearly sets our framework apart from other FCM analysis
tools. By factoring out as much of the bookkeeping as possible,
programmers can focus on the actual operations rather then having to
deal with the tedious details of data integration and
access. Third-party methods can act on their own as first-class
citizens in the analysis framework, without breaking the work flow or
the basic infrastructure. This design allows for the straightforward
extension of \Rpackage{flowCore}'s capabilities, and has already
fostered the development of a number of valuable add-ons
\cite{lo2008agf,sarkar2008ufv}.

\subsubsection*{Transformation and compensation}

Data transformation is essential for both data visualization and
modelling \cite{lo2008agf}. The major transformations that are
routinely used in FCM analysis have been implemented in
\Rpackage{flowCore} (\textit{e.g.,} log, bi-exponantioal or arcsinh,
see Table~\ref{table1} for a complete list). Furthermore, the design
of the R language makes it easy to define arbitrary functions to apply
to the data of individual \Rclass{flowFrames} or entire
\Rclass{flowSets}, respectively. Compensation is available for both
\Rclass{flowFrames} and \Rclass{flowSets}, and the software also
offers functionality to compute spillover or compensation matrices
from a set of appropriate single stain controls.

\subsubsection*{Gating}
In \Rpackage{flowCore}, gating operations are represented by classes
that can be extended in an object-oriented manner
(Table~\ref{table2}). Basic gate types such as rectangular gates,
ellipses and polygon gates are implemented as part of the
framework. In addition, we introduce the notion of data-driven gates,
or filters, for which the necessary parameters are computed based on
the properties of the underlying data, for instance by modelling data
distribution or by density estimation. This approach is fundamentally
different from the traditional application of static gating regions
across samples, as it is able to take into accounts unforeseen changes
in signal intensities, such as drifts in the instrumentation
over time or sample variablity.

The ability to programmatically access gates is a prerequisite for
automated or semi-automated gating. By utilizing a unified interface
for all different types of gates, the user is able to subset data sets
as well as to create summary statistics, for example the proportions
of events falling in a single gate or in a combination of
gates. Complex combinations and hierarchies of gates can be captured
in objects of class \Rclass{filterSet}, allowing to apply multi-step
gating strategies. The definition of gates in \Rpackage{flowCore}
follows the Gating-ML CR, thus any \Rpackage{flowCore} gating strategy
can be reproduced by any other software that also adheres to the
standard.

Gating, as well as all other operations in \Rclass{flowCore}, can be
applied over each individual frame in a \Rclass{flowSet}, and summary
methods provide information about the outcome of these operations. In
addition, the result of a gating operation can be used to subset the
input \Rclass{flowFrame} or \Rclass{flowSet}, either by filtering out
negative events or by splitting in multiple subpopulations. This
design allows to easily combine all of \Rpackage{flowCore}'s
components into complex work flows.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results and Discussion %%
%%
\section*{Results and Discussion}
  \subsection*{From quality assessment to batch gating}
  The \Rpackage{flowCore} package has been successfully applied in the
  analysis of several datasets
  \cite{gasparetto2004ice,brinkman2007hcf}. A complete documentation
  of the \Rpackage{flowCore} software is beyond the scope of this
  publication.  Much more comprehensive documentation and users guide
  information with programmatic examples is available online
  (\url{http://bioconductor.org/packages/2.2/bioc/html/flowCore.html})
  and as part of the package distribution. Here, we want to briefly
  exemplify some of the software's key features, that is, the coherent
  treatment of all samples in a potentially large experiment, the
  concept of data-driven automated gating, the integration of existing
  software into the framework, and the generation of
  publication-quality graphics for data visualization.

  Data analysis for most experiments usually begins with a quality
  assurance step, and we can use functionality from the
  \Rpackage{flowQ} package to create an HTML report that highlights
  potential quality issues. Assuming that the data has already been
  imported as the \Rclass{flowSet} object ``dat'', the following
  simple line of code produces the output shown in Figure~\ref{flowQ}:

\begin{verbatim}
> library(flowCore)
> library(flowQ)
> qaReport(dat, c("qaProcess.timeline", "qaProcess.timeflow", 
                  "qaProcess.cellnumber"))
\end{verbatim}

  The report is interactive and provides drill-down to more detailed
  aspects of the analysis, starting from a concise overview. The
  design of \Rpackage{flowCore}'s data model allows for a coherent
  treatment of all the samples, hence we are able to compare features
  between individuals, or between groups of individuals, based on the
  available metadata information.

  Static gating for all samples in a high-throughput FCM experiment is
  often impossible, since the measured variables tend to vary between
  different treatments, over time or between different experiment
  batches. Automated or data-driven gating has the potential to
  estimate the gating regions from the underlying data, thus providing
  a fast objective solution to the analysis of potentially very large
  and diverse data sets \cite{lo2008agf}. One of the automated gating
  methods implemented in \Rpackage{flowCore} is based on identifying
  areas of significant curvature in a kernel density estimate of the
  data \cite{wand2008}. Assuming that the regions of interest are of
  high density, the software is able to reliably detect them in a one
  or two-dimensional density landscape.

\begin{verbatim}
> cf <- curv2Filter("FL1-H", "FL3-H")
> fres <- filter(dat, cf)
\end{verbatim}

Kernel density estimation is a well-known problem in statistical
computing, and a lot of effort has been invested in the development of
good software to address it. The modular design of \Rpackage{flowCore}
allows to easily integrate these existing solutions into our
framework. In this example, we directly use R code from the
\Rpackage{feature} package developed by \cite{wand2008}. Instead of
re-writing existing code, we are able to include it via the well
tested distribution mechanism provided by R's software package
system. This process is bi-directional, and all functionality
implemented in \Rpackage{flowCore} is available to other package
authors, as we have seen with the afore-mentioned \Rpackage{flowClust}
package.

We can chose one of the many visualization option from the flowViz
package to plot the results of the recent filtering operation. A very
basic matrix of density plots is shown in Figure \ref{xyplot}, where
each panel in the matrix represents the data of one individual
patient.

\begin{verbatim}
> xyplot(`FL1-H` ~ `FL3-H` | SampleID, data=dat, filter=fres)
\end{verbatim}

\subsection*{Workflow proposal(??)}

Florian?

\subsection*{Related flow packages}

In addition to the \Rpackage{flowCore} package that offers basic
infrastructure, we have implemented a range of additional Bioconductor
packages that are dedicated to more specific tasks of FCM data
analysis. The \Rpackage{flowViz} package \cite{sarkar2008ufv}
provides sophisticated data visualization tools, that make use off
multivariate trellis plotting \cite{lattice}.  These functions can be
used to quickly generate customized plots for extended cytometry data
sets for both direct data inspection and quality control.  The objects
metadata information can be used to arrange the layout and composition
of the plots.  Furthermore, the design and the API of the
visualization software is very generic, and users can readily extend
its capabilities by providing self-defined plotting functions.  The
\Rpackage{flowQ} package offers more advanced quality assurance (QA)
methodology and a framework to create interactive web-based reports of
QA results. The \Rpackage{flowUtil} package implements data import and
export including flow-cytometry specific standard markup
language. Finally, the \Rpackage{flowStats} package provides elaborate
statistical methods that are relevant in the context of flow
cytometry data analysis. \cite{lo2008agf} have recently developed an
automatic gating approach via robust model-based clustering using
\Rpackage{flowCore}'s data model and infrastructure which is
implemented in the Bioconductor package \Rpackage{flowClust}. Another
package, \Rpackage{plateCore}, providing more specialized support for
experiments conducted on microtitre plates and facilitiating the
handling of spatial metadata, is under development.

%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusions}
Through \Rpackage{flowCore}, we have provided the FCM community with
an open source, freely available, highly functional, standards
compliant development and analysis platform for high throughput data
analysis.  Our intention is to provide a platform for the
collaborative development of new analysis methods that will facilitate
the transition of these new methods to the larger flow community.  Our
experience has been that the collaborative effort we enable for
devising new methodology has been proven beneficial for a number of
different biological and computational biology challenges.  We hope
that our framework will be the foundation for fruitful shared research
by many collaborators from multiple scientific fields and will help
resolve bottlenecks that currently prevent further development and
deployment of FC-HCS to increasingly complex and important scientific
and clinical applications.

%%%%%%%%%%%%%%%%%%%%%%
\section*{Availability and requirements}
Project name: flowCore
Project home page: http://bioconductor.org 
Operating system(s): The flowCore package can be installed on all the 
platforms supporting R. These include a wide variety of UNIX platforms, Windows 
and MacOS.  
Programming language: R 
Licence: The Artistic License, Version 2.0

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}
The project was conceived by RB, RG, PH. It was designed, and
developed by all the authors.  NLM, FH, DS, and BE wrote most of the
software’s code. The manuscript was prepared by FH and NLM, then
revised and approved by all authors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgements}
  This work was supported by NIH grant EB005034 and by the Michael Smith
Foundation for Health Research. RRB is an ISAC Scholar.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%              
%%  Bmc_article.bst  will be used to                       %%
%%  create a .BBL file for submission, which includes      %%
%%  XML structured for BMC.                                %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %% 
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %% 
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{flowCoreRef} }     % Bibliography file (usually '*.bib' )

%%%%%%%%%%%

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}
  \subsection*{Figure 1 - flowCore Framework}
  \begin{figure}[hp]
    \centering
    \includegraphics[width=0.9\textwidth]{Figure1-flowCoreFrameWork.jpg}
    \caption{\label{fig1:FrameWork}{For each experiment, the content of
        the FCS files, phenotypic and metadata are stored in a
        \Rclass{flowSet}. Each \Rclass{flowFrame} in a \Rclass{flowSet}
        corresponds to one FCS file. All basic operations (e.g.,
        compensation, transformation, gating) can be applied to either
        single \Rclass{flowFrames} or a \Rclass{flowSet} simultaneously.}}
  \end{figure}

  \subsection*{Figure 2 - Quality assessment}
    \begin{figure}[hp]
      \centering
      \includegraphics[width=0.98\textwidth]{flowQ.jpg}
      \caption{\label{flowQ}%
        HTML quality assessment report generated by the flowQ package for a
        subset of data from an experiment focussing on Graft-Versus-Host
        Disease \citep{brinkman2007hcf}. Rows correspond to the samples in
        the set, columns to different quality checks.}
    \end{figure}

    \subsection*{Figure 3 - Quality assessment}
    \begin{figure}[hp]
      \centering
      \includegraphics[width=0.75\textwidth]{xyplot}
      \caption{\label{xyplot}%
        Scatterplot matrix of a single \Rclass{flowSet} from an experiement
        focussing on immune tolerance following kidney
        transplantation. Outlines of the gating regions identified by an
        automated gating operation are added on top of the density
        representation of the data.}
    \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
  \subsection*{Table 1 - Data transformations}
    \begin{table}[ht]
      \caption{\label{table1} Within these formulas, $x$ is the variable
          corresponding to value being transformed, $y$ is the result of the
          transformation, $a$, $b$, $c$, $d$, $f$, $p$, $m$, $T$, and $w$, are
          constants affecting the transformation function, $e$ is the base of
          the natural logarithm, and $root$ is a root-finding algorithm.  }
      \begin{center}
        \begin{tabular}{|l|l|}
          \hline
          \multicolumn{2}{|c|}{Data Transformations} \\
          \hline
          linear & $ax + b$ \\
          quadratic & $ax^2 + bx + c$ \\
          natural logarithm & $ln(x)(r/d)$ \\
          logarithm & $log(x)(r/d)$ \\
          biexponential & $root(a*e^{(b*y)}-c*e^{(-d+y)}+f)$ \\
          logicle& $root(Te^{-(m-w)}*(e^{(y-w)}-p^2e^{-(y-w)/p}+p^2-1)$ \\
          truncate & $x_{x<=a} = a$ \\
          scale & $(x-a)/(b-a)$ \\
          split-scale & combination of linear and logarithm \\
          arcsinh & $asinh(a + bx)+c$ \\
          \hline
        \end{tabular}
      \end{center}
    \end{table}

  \subsection*{Table 2 - Filter and gate classes}
  \begin{table}[ht]
    \begin{center}
      \begin{tabular}{|l|l|}
        \hline
        \multicolumn{2}{|c|}{Gates} \\
        \hline
        rectangleGate & n-dimensional rectangular regions \\
        quadGate & quadrant regions in two dimensions \\
        polygonGate & polygonal regions in two dimensions \\
        polytopeGate & generalization of polygon in n dimensions \\
        ellipsoidGate & n-dimensional ellipsoid region \\
        \hline
        \multicolumn{2}{|c|}{Filters} \\
        \hline
        sampleFilter & random sub-sampling\\
        expressionFilter & results of a boolean expression \\
        kmeansFilter & K-means clustering \\
        norm2Filter & bivariate normal distribution \\
        curv1Filter & local density regions in 1D \\
        curv2Filter & density regions in 2D \\
        timeFilter & abnormal data acquisition over time \\
        \hline
        filterSet & gating strategies \\
        \hline
      \end{tabular}
    \end{center}
  \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{bmcformat}
\end{document}